### 2017-05-04 paper.Rnw
Add paper template

\documentclass{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,float}
\usepackage{multirow,setspace}
\usepackage{natbib,enumerate}
\usepackage{booktabs}
\usepackage{placeins}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\tab}{\hspace{0.5cm}}
\newcommand{\modref}[1]{(\ref{#1})}
\mathchardef\mhyphen="2D

\newcounter{DefnCounter}
\setcounter{DefnCounter}{1}

\newcounter{ThmCounter}
\setcounter{ThmCounter}{1}

\newcounter{ExampleCounter}
\setcounter{ExampleCounter}{1}

\newcommand{\defn}[1]{\textsc{Definition 1.\arabic{DefnCounter}\stepcounter{DefnCounter}: #1\\} }
\newcommand{\thm}{\textsc{Theorem 1.\arabic{ThmCounter}\stepcounter{ThmCounter}\\} }
\newcommand{\ex}{\textsc{Example 1.\arabic{ExampleCounter}\stepcounter{ExampleCounter}\\} }

\newcommand{\Defn}{\underline{Definition}}
\newcommand{\Q}{\underline{Question}}
\newcommand{\Qs}{\underline{Questions}}
\newcommand{\bbeta}{{\mbox{\boldmath$\beta$}}}
\newcommand{\bmu}{{\mbox{\boldmath$\mu$}}}
\newcommand{\balpha}{{\mbox{\boldmath$\alpha$}}}
\newcommand{\btheta}{{\mbox{\boldmath$\theta$}}}
\newcommand{\bphi}{{\mbox{\boldmath$\phi$}}}
\newcommand{\bSigma}{{\mbox{\boldmath$\Sigma$}}}
\newcommand{\bLambda}{{\mbox{\boldmath$\Lambda$}}}
\newcommand{\bpi}{{\mbox{\boldmath$\pi$}}}
\newcommand{\R}{\texttt{R}}
\newcommand{\Lik}{\mathcal{L}}
\newcommand{\bx}{\textbf{x}}
\newcommand{\by}{\textbf{y}}
\newcommand{\bX}{\textbf{X}}
\newcommand{\sic}{\text{Inv-}\chi^2}

\newcommand{\sao}{SaO$_2$}

\setlength{\marginparwidth}{2cm}
\begin{document}
\SweaveOpts{concordance=TRUE}
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of medical measurements about severe weather in the United States.

The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models and factors that will be used in the analysis. Section~\ref{s:results} describes the results of the analysis. And Section~\ref{s:discussion} contains the conclusions.

\section{Results}
\label{s:results}

\section{Methods}
\label{s:methods}

\section{Discussion}
\label{s:discussion}

\clearpage
\section*{Appendix A}
This appendix includes the code used to implement the models.

\section*{Appendix B}
This appendix will include details on the calculations required to derive our models.
\end{document}

### 2017-05-06 paper.Rnw
Flesh out template, add references, and first draft of introduction

@@
 \usepackage{multirow,setspace}
-\usepackage{natbib,enumerate}
+\usepackage{enumerate}
 \usepackage{booktabs}
@@
 \label{s:intro}
-This analysis will be done on a data set of a variety of medical measurements about severe weather in the United States.
+This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
+The data set contains a variety of measures from severe weather events in the United States from 1996-2016. 
+For this project we focused on the deaths directly attributable to the event.
+Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
+In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.
 
-The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models and factors that will be used in the analysis.
+The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models.
 Section~\ref{s:results} describes the results of the analysis.
@@
 
+\section{Methods}
+\label{s:methods}
+The deaths attributed to a severe weather event is `count' data. 
+The most common model used for count data is the Poisson distribution. 
+However for some weather events, the negative binomial model is a better fit, because the Poisson distribution assumes that the events being counted occur independently.
+
+Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants.
+These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$.
+
+We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.
+
+\subsection{Poisson}
+\label{ss:mPoisson}
+
+\subsection{Negative Binomial}
+\label{ss:mNBinom}
+
+\subsection{Zero Inflated Poisson}
+\label{ss:mZiPoisson}
+
+\subsection{Zero Inflated Negative Binomial}
+\label{ss:mZiNBinom}
+
 \section{Results}
 \label{s:results}
+\subsection{Poisson}
+\label{ss:rPoisson}
 
-\section{Methods}
-\label{s:methods}
+\subsection{Negative Binomial}
+\label{ss:rNBinom}
+
+\subsection{Zero Inflated Poisson}
+\label{ss:rZiPoisson}
+
+\subsection{Zero Inflated Negative Binomial}
+\label{ss:rZiNBinom}
 
@@
 
+\begin{thebibliography}{1}
+  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
+    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
+    Accessed April 2017.
+\end{thebibliography}
 \clearpage

### 2017-05-06 paper.tex
Add paper template
Flesh out template, add references, and first draft of introduction

--- 2017-05-04 paper.Rnw
@@
 \setlength{\marginparwidth}{2cm}
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{paper-concordance}
 \begin{center}

### 2017-05-07 paper.Rnw
ZINB derivation

@@
 \label{ss:mZiNBinom}
+The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
+For a single $X$ the probability density is: 
+$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}.$$
+We take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. For a full derivation, see \ref{a:dZINB}. My posterior is:
+$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)$$
+This distribution does not factor nicely, so I will use the Metropolis algorithm to sample from it. 
+Because this posterior does not suggest any obvious proposal distributions I will sample each independently from a normal distribution centered at $\theta^*$, and with a variance that is tuned to yeild an appropriate acceptance rate.
 
@@
 \clearpage
-\section*{Appendix A}
+\appendix
+\section{Code}
+\label{a:code}
 This appendix includes the code used to implement the models.
 
-\section*{Appendix B}
+\section{Derivations}
+\label{a:derivation}
 This appendix will include details on the calculations required to derive our models.
+\subsection{Zero Inflated Negative Binomial Derivation}
+\label{a:dZINB}
+The likelihood for the ZINB is 
+\begin{align*}
+\mathcal{L}(X|\sigma, p, r) &= \prod_{i=1}^N \sigma I_{X=0}(X_i) + (1-\sigma)\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\\
+\intertext{For ease of notation let $Z$ be the number of zero values in $X$, and $N$ be the total number of observations.}
+&=\prod_{X_i=0}\left(\sigma + (1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
+&=\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
+&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
+&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^n\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)
+\end{align*}
+As mentioned in section \ref{ss:mZiNBinom} we take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. Therefore my joint posterior is:
+\begin{align*}
+p(r,\sigma, p|X)\propto&\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
+\intertext{I am now going to take the log of the posterior because it helps with computation}
+\ln\left(p(r,\sigma, p|X)\right)\propto&Z\ln\left(1 + (1/\sigma-1)(1-p)^r\right) + Z\ln(\sigma) + (N-Z)\ln(1-\sigma)+(N-Z)r\ln(1-p)\\
+&+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
+%\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
+\end{align*}
+
 \end{document}

### 2017-05-08 paper.tex
Added ZIP Methods

@@
 \label{ss:mZiPoisson}
+The Zero Inflated Poisson (ZIP) model has two parameters.
+The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
+For a single observation x, the probability density is:
+
+\[
+p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
+\]
+
+\noindent We can write the likelihood as follows:
+$$
+L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
+$$
+
+\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+
+\[
+\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
+\]
+
+\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
+
+\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
+
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
+
+\noindent Neither of these distributions is recognizable.
+We can use a Metropolis-Hastings algorithm to sample from both of them.
+We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
+We will tune them to obtain a better acceptance rate.
 
@@
     Accessed April 2017.
+   \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
 \end{thebibliography}

### 2017-05-09 zinb_sampler.R
Merge remote-tracking branch 'origin/master'
# Conflicts:
#	paper.pdf

@@ -1,0 +1,134 @@
+logpost <- function(sigma, p, r, Z, N, S, nonzero){
+  lp <- Z*log(1 + (1/sigma-1)*(1-p)^r)
+  lp <- lp + Z*log(sigma)
+  lp <- lp + (N-Z)*log(1-sigma)
+  lp <- lp + (N-Z)*r*log(1-p)
+  lp <- lp + S*log(p)
+  lp <- lp - log(r)/2
+  lp <- lp - (N-Z)*lgamma(r)
+  lp <- lp + sum(lgamma(r+nonzero))
+  return(lp)
+}
+
+sample <- function(r.mean, p.mean, sigma.mean, r.var, p.var, sigma.var){
+  sigma.star <- rtruncnorm(1, 0, 1, sigma.mean, sqrt(sigma.var))
+  #rbeta(1, sigma.mean*sigma.N, (1-sigma.mean)*sigma.N)
+  p.star <- rtruncnorm(1, 0, 1, p.mean, sqrt(p.var))
+  #rbeta(1, p.mean*p.N, (1-p.mean)*p.N)
+  r.star <- rtruncnorm(1, 0, Inf, r.mean, sqrt(r.var))
+  return(c(sigma.star, p.star, r.star))
+}
+
+g <-function(sigma, p, r, 
+             sigma.mean, p.mean, r.mean, 
+             sigma.var, p.var, r.var){
+    return(dtruncnorm(sigma, 0, 1, sigma.mean, sqrt(sigma.var))*
+           dtruncnorm(p, 0, 1, p.mean, sqrt(p.var))*
+           dtruncnorm(r, 0, Inf, r.mean, sqrt(r.var)))
+}
+
+
+zinb.sampler <- function(df, year, event_type, chainlen, 
+                         r.var=1, p.var=1, sigma.var=1, 
+                         burnin=0, thinning=1, nchains=1, random=TRUE){
+  X <- df[df$YEAR>year&df$EVENT_TYPE==event_type,"DEATHS_DIRECT"]
+  N <- length(X)
+  Z <- sum(X==0)
+  S <- sum(X)
+  nonzero <- X[X!=0]
+  B <- chainlen
+  b <- burnin
+  
+  r.chains <- c()
+  sigma.chains <- c()
+  p.chains <- c()
+  ar <- c()
+  
+  for(j in 1:nchains){
+    r.array <- rep(0, B)
+    sigma.array <- rep(0, B)
[…]

### 2017-05-09 paper.tex
Merge remote-tracking branch 'origin/master'
# Conflicts:
#	paper.pdf
Move new content from paper.tex to paper.Rnw

@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[margin=0.75in]{geometry}
@@
 \usepackage{placeins}
+\usepackage{cite}
 \usepackage{caption}
@@
 \setlength{\marginparwidth}{2cm}
-\usepackage{Sweave}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-\input{paper-concordance}
 \begin{center}
@@
 This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
-The data set contains a variety of measures from severe weather events in the United States from 1996-2016. 
+The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
 For this project we focused on the deaths directly attributable to the event.
@@
 
-\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
 
@@
 
-\begin{thebibliography}{1}
+\begin{thebibliography}{2}
   \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
@@
 &+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
-%\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
 \end{align*}

### 2017-05-09 paper.Rnw
Merge remote-tracking branch 'origin/master'
# Conflicts:
#	paper.pdf

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 \begin{center}
@@
 \label{ss:mZiPoisson}
+The Zero Inflated Poisson (ZIP) model has two parameters.
+The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
+For a single observation x, the probability density is:
+
+\[
+p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
+\]
+
+\noindent We can write the likelihood as follows:
+$$
+L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
+$$
+
+\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+
+\[
+\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
+\]
+
+\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
+
+\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
+
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
+
+\noindent Neither of these distributions is recognizable.
+We can use a Metropolis-Hastings algorithm to sample from both of them.
+We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
+We will tune them to obtain a better acceptance rate.
 
@@
     Accessed April 2017.
+  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
 \end{thebibliography}

### 2017-05-09 paper.Rnw
trying to fix sweave
added poisson and negative binomial models to paper.Rnw

--- 2017-05-07 paper.Rnw
@@
 \label{ss:mPoisson}
+The Poisson model has one parameter, $\lambda$ represents the expected number of occurances of the event of interest. For a single random variable $x$, the probability density is:
+$$p(x|\lambda)=\frac{\lambda^xe^{-\lambda}}{x!}.$$
+Thus the likelihood for a sum of Poisson random variables can be written as follows:
+$$\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{X}}e^{-n\lambda}.$$ 
+Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
+$$\pi(\lambda)\propto\lambda^{1/2-1}e^{-0\cdot\lambda}.$$
+We recognize this as the kernel of an improper gamma distribution. Combining the likelihood and the prior distribtion yields the following posterior distribution:
+$$p(\lambda|X)=\lambda^{n\bar{X}+1/2-1}e^{-n\lambda}.$$
+We recognize this as the kernel of a gamma distribution, namely
+$$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$
 
@@
 \label{ss:mNBinom}
+The Negative Binomial model has two parameters, $r,p$ represents the expected number of occurances of the event of interest. For a single random variable $x$, the probability density is:
+$$p(X|r,p)=\frac{\Gamma(r+x)}{\Gamma(r)x!}p^{x}(1-p)^{r}.$$
+
+Thus the likelihood for a sum of Negative Binomial random variables can be written as follows:
+$$\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}.$$
+Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
+$$\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}.$$
+Combining the likelihood and the prior distribtion yields the following posterior distribution:
+$$p(r,p|X)=\Bigg\{\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}\Bigg\}\Bigg\{r^{1/2}p^{-1}(1-2)^{-1/2}\Bigg\}$$
+From this posterior we obtain the full conditionals. First consider $p|r,X$:
+$$p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$$
+We recognize this as the kernel of a beta distribution, namely
+
+$$p|r,X\sim\mathcal{B}eta(n\bar{X},nr+1/2).$$
+Next consider $r|p,X$:
+$$p(r|p,X)\propto\bigg[\prod_{i=1}^n\Gamma(r+x_i)\bigg]\Gamma(r)^{-n}(1-p)^{nr}r^{1/2}$$
+This is not a recognized distribtion. So if we wish to make inferences on $r$ we must use a Metropolis algorithm to sample from it.
 

### 2017-05-09 paper.tex
added poisson and negative binomial models to paper.Rnw

--- 2017-05-08 paper.tex
@@
 \label{ss:mZiPoisson}
-The Zero Inflated Poisson (ZIP) model has two parameters.
-The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
-For a single observation x, the probability density is:
-
-\[
-p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
-\]
-
-\noindent We can write the likelihood as follows:
-$$
-L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
-$$
-
-\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
-
-\[
-\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
-\]
-
-\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
-
-\[
-p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
-\]
-
-\[
-p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
-\]
-
-\noindent Neither of these distributions is recognizable.
-We can use a Metropolis-Hastings algorithm to sample from both of them.
-We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
-We will tune them to obtain a better acceptance rate.
 
@@
     Accessed April 2017.
-   \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
 \end{thebibliography}

### 2017-05-09 clean storm data.Rmd
cleaned storm data

---
title: "R Notebook"
output: html_notebook
---
```{r warning = FALSE, message = FALSE, echo = FALSE}
setwd("/users/nikopaulson/documents/homework/georgetown/math640")
library(tidyverse)

storm <- read_csv("storm_details.csv")

events_from_codebook <- c("Astronomical Low Tide", "Avalanche", "Blizzard", "Coastal Flood", "Cold/Wind Chill", "Debris Flow", "Dense Fog", "Dense Smoke", "Drought", "Dust Devil", "Dust Storm", "Excessive Heat", "Extreme Cold/Wind Chill", "Flash Flood", "Flood", "Freezing Fog", "Frost/Freeze", "Funnel Cloud", "Hail", "Heat", "Heavy Rain", "Heavy Snow", "High Surf", "High Wind", "Hurricane/Typhoon", "Ice Storm", "Lakeshore Flood", "Lake-Effect Snow", "Lightning", "Marine Hail", "Marine High Wind", "Marine Strong Wind", "Marine Thunderstorm Wind", "Rip Current", "Seiche", "Sleet", "Storm Tide", "Strong Wind", "Thunderstorm Wind", "Tornado", "Tropical Depression", "Tropical Storm", "Tsunami", "Volcanic Ash", "Waterspout", "Wildfire", "Winter Storm", "Winter Weather")

have <- toupper(sort(unique(storm$EVENT_TYPE)))
want <- toupper(events_from_codebook)
must_rename <- have[which(have %in% want == FALSE)]

df <- storm %>%
  mutate(EVENT_TYPE = toupper(EVENT_TYPE)) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "HAIL FLOODING", "FLOOD")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "FLOODING", "FLOOD")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "HAIL/ICY ROADS", "HAIL")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "HEAVY WIND", "HIGH WIND")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "HIGH SNOW", "HEAVY SNOW")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE %in% c("HURRICANE",
                                                            "HURRICANE (TYPHOON)"), "HURRICANE/TYPHOON")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "LANDSLIDE", "DEBRIS FLOW")) %>%         
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "MARINE DENSE FOG", "DENSE FOG")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "MARINE HURRICANE/TYPHOON", "HURRICANE/TYPHOON")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "MARINE LIGHTNING", "LIGHTNING")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "MARINE TROPICAL DEPRESSION", "TROPICAL DEPRESSION")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "MARINE TROPICAL STORM", "TROPICAL STORM")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "SNEAKERWAVE", "TSUNAMI")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "STORM SURGE/TIDE", "STORM TIDE")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE %in% c("THUNDERSTORM WIND/ TREE", 
                                                            "THUNDERSTORM WIND/ TREES", 
                                                            "THUNDERSTORM WINDS FUNNEL CLOU", 
                                                            "THUNDERSTORM WINDS HEAVY RAIN", 
                                                            "THUNDERSTORM WINDS LIGHTNING", 
                                                            "THUNDERSTORM WINDS/HEAVY RAIN", 
                                                            "THUNDERSTORM WINDS/ FLOOD", 
                                                            "THUNDERSTORM WINDS/FLASH FLOOD", 
                                                            "THUNDERSTORM WINDS/FLOODING"), "THUNDERSTORM WIND")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE %in% c("TORNADO/WATERSPOUT",
                                                            "TORNADOES, TSTM WIND, HAIL"), "TORNADO")) %>%
  mutate(EVENT_TYPE = replace(EVENT_TYPE, EVENT_TYPE == "VOLCANIC ASHFALL", "VOLCANIC ASH")) %>%
  filter(! (EVENT_TYPE == "NORTHERN LIGHTS" | EVENT_TYPE == "OTHER")) %>%
  filter(YEAR %in% 1996:2016) %>%
  select(EVENT_ID, YEAR, STATE, EVENT_TYPE, DEATHS_DIRECT)

summaryData <- df %>%
  group_by(EVENT_TYPE) %>%
  summarize(count = n(), fatalities = sum(DEATHS_DIRECT)) %>%
  arrange(desc(fatalities)) %>%
  select(EVENT_TYPE, fatalities)

ggplot(summaryData, aes(x = reorder(EVENT_TYPE, - fatalities), y = fatalities)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Deadliest Weather Events: 1996-2016", 
       x = "Event Type", 
       y = "Fatalities") + 
  theme(axis.text.x = element_text(angle = - 90, hjust = 0))
```

### 2017-05-09 clean storm data.Rmd
R code for Poisson and NB posterior inference

@@
 title: "R Notebook"
-output: html_notebook
+output:
+  pdf_document: default
+  html_notebook: default
 ---
@@
 ```
+
+```{r echo = FALSE}
+summaryData2 <- df %>%
+  filter(EVENT_TYPE == "TORNADO") %>%
+  group_by(YEAR) %>%
+  summarize(TOR_COUNT = n(), DEATHS = sum(DEATHS_DIRECT)) 
+```
+
+#POISSON
+
+Likelihood:      $\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{x}}e^{-n\lambda}$
+
+Jeffreys' Prior: $\pi(\lambda)=\lambda^{1/2-1}e^{-0\cdot\lambda}$
+
+Posterior:       $p(\lambda|X)=\lambda^{n\bar{x}+1/2-1}e^{-n\lambda}$
+
+So the posterior on lambda with non-informative prior is $\mathcal{G}amma(n\bar{x}+1/2,n)$.
+
+```{r}
+set.seed(05092017)
+
+df_tor <- df %>%
+  filter(EVENT_TYPE == "TORNADO")
+
+X <- df_tor$DEATHS_DIRECT
+alpha <- sum(X) + 1 / 2
+beta <- length(X)
+B <- 20000
+
+post_lambda <- rgamma(B, alpha, beta)
+ci_lambda <- quantile(post_lambda, c(0.025, 0.975))
+```
+
+Using the data based off of 20000 posterior samples, we get the posterior median for $\lambda$ is `r round(median(post_lambda), 4)` with a 95% credible interval of `r round(ci_lambda[1], 4)` to `r round(ci_lambda[2], 4)`.
+The posterior denisty is in the figure below:
+
+```{r echo = FALSE}
+plot(density(post_lambda),
+     xlab = expression(lambda),
+     ylab = expression(paste("p(", lambda, "|x)")),
+     main = "Density plot with 95% credible interval", 
+     lwd = 2,
+     col = "blue")
+credInt <- quantile(post_lambda, probs = c(0.025, 0.975)) 
+densOut <- density(post_lambda)
+idStart <- max(which(densOut$x < credInt[1])) + 1 
+idEnd <- min(which(densOut$x > credInt[2])) - 1
+gx <- densOut$x[idStart : idEnd]
+gy <- densOut$y[idStart : idEnd]
+px <- rep(0, length(gy))
+polygon(c(gx, rev(gx)), c(px, rev(gy)), border = FALSE, col = rgb(0, 0, 1, alpha = 0.5)) 
+abline(h = 0)
+```
+
+#NEGATIVE BINOMIAL
+
+Likelihood: $\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{x}}(1-p)^{nr}$
+
+Jeffreys' Prior: $\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}$
+
+Full Posterior: $p(r,p|X)=\mathcal{L}(X|r,p)\pi(r,p)$
+
+Conditional of p: $p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$
+
+Conditional of r: $p(r|p,X)\propto\big[\prod_{i=1}^n\Gamma(r+x_i)\big]\Gamma(r)^{-n}(1-p)^{nr}r^{1/2}$
+
+```{r}
+set.seed(05092017)
+
+X <- df_tor$DEATHS_DIRECT
+
+r <- 1 
+alpha <- sum(X)
+beta <- length(X) * r + 1 / 2
+B <- 20000
+
+post_p <- rbeta(B, alpha, beta)
+ci_p <- quantile(post_p, c(0.025, 0.975))
+```
+
+Using the data based off of 20000 posterior samples, we get the posterior median for $p$ is `r round(median(post_p), 4)` with a 95% credible interval of `r round(ci_p[1], 4)` to `r round(ci_p[2], 4)`.
+The posterior denisty is in the figure below:
+
+```{r echo = FALSE}
+plot(density(post_p),
+     xlab = expression(p),
+     ylab = expression(paste("p(", p, "|r,x)")),
+     main = "Density plot with 95% credible interval", 
+     lwd = 2,
+     col = "blue")
+credInt <- quantile(post_p, probs = c(0.025, 0.975)) 
+densOut <- density(post_p)
+idStart <- max(which(densOut$x < credInt[1])) + 1 
+idEnd <- min(which(densOut$x > credInt[2])) - 1
+gx <- densOut$x[idStart : idEnd]
+gy <- densOut$y[idStart : idEnd]
+px <- rep(0, length(gy))
+polygon(c(gx, rev(gx)), c(px, rev(gy)), border = FALSE, col = rgb(0, 0, 1, alpha = 0.5)) 
+abline(h = 0)
+```

### 2017-05-10 paper.Rnw
ZIP Methods and tentative Results

@@
 
+The Zero Inflated Poisson (ZIP) model has two parameters.
+The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
+For a single observation x, the probability density is:
+
+\[
+p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
+\]
+
+\noindent We can write the likelihood as follows:
+$$
+L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
+$$
+
+\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+
+\[
+\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
+\]
+
+\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
+
+\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
+
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
+
+\noindent Neither of these distributions is recognizable.
+We can use a Metropolis-Hastings algorithm to sample from both of them.
+We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
+We will tune them to obtain a better acceptance rate.
+
 \subsection{Zero Inflated Negative Binomial}
@@
 
+In tuning the parameters of the two models, slightly different values were chosen for the two different event types.
+The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods.
+The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model.
+
+For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are as follows:
+
+\[\]
+\noindent Table 1: Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types.
+The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
+The third does the same for p.
+All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.
+
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$ & p  \\ \hline
+    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
+    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
+    \end{tabular}
+\end{center}
+
 \subsection{Zero Inflated Negative Binomial}

### 2017-05-10 paper.Rnw
finish zing results

--- 2017-05-09 paper.Rnw
@@
 \usepackage{wrapfig}
-\usepackage{xcolor}
 \usepackage{hyperref}
@@
 \begin{document}
+<<imports, cache=FALSE, echo=FALSE>>= 
+read_chunk('zinb_sampler.R')
+options(scipen=4, digits=3)
+library('truncnorm')
+@
+<<dataset, echo=FALSE, cache=FALSE>>= 
+load('small_storms.RData')
+storm_details <- small_storms
+@
 \begin{center}
@@
 \label{ss:rZiNBinom}
+<<zinbsampler, echo=FALSE>>=
+@
+<<densityplot, echo=FALSE>>=
+@
+<<zinbparams, echo=FALSE, cache=TRUE>>=
+samplesize <- 650000
+burnin <- 20000
+thin <- 70
+r.var=.00002
+p.var=.00001
+sigma.var=.001
+@
+The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution.
+For the source code of the sampler, please see \ref{a:cZINB}.
 
+This model suffered severely from the curse of dimensionality.
+It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary.
+Autocorrelation and running mean plots for both the Tornado and the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+<<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
+flood.zinb <- zinb.sampler(storm_details, "Flash Flood", samplesize, 
+                           r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
+                           burnin=burnin, thinning=thin)
+@
+<<tornadosample, cache=TRUE, cache.vars='tornado.zinb', echo=FALSE, dependson='zinbparams'>>=
+tornado.zinb <- zinb.sampler(storm_details, "Tornado", samplesize, 
+                             r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
+                             burnin=burnin, thinning=thin)
+@
+In the end $\Sexpr{samplesize}$ samples were drawn originally for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
+The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$.
+The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
+<<evaluation, echo=FALSE, cache=FALSE>>=
+@
+<<model-eval, echo=FALSE, cache=TRUE, dependson=c('floodsample', 'tornadosample')>>=
+flood.dic.zinb <- dic(storm_details, "Flash Flood", flood.zinb$sigma, 
+                      flood.zinb$r, flood.zinb$p)
+tornado.dic.zinb <- dic(storm_details, "Tornado", tornado.zinb$sigma, 
+                        tornado.zinb$r, tornado.zinb$p)
+#flood.wiac.zinb <- wiac(storm_details, "Flash Flood", flood.zinb$sigma, 
+#                   flood.zinb$r, flood.zinb$p)
+#tornado.wiac.zinb <- wiac(storm_details, "Tornado", tornado.zinb$sigma, 
+#                          tornado.zinb$r, tornado.zinb$p)
+@
+
+The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.
+I was unable to calculate WIACs because it was too computationally expensive.
+\begin{table}
+\centering
+\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
+\label{t:ZINBparams}
+\begin{tabular}{lcccccc}
+\toprule
+&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
+\cmidrule(r){2-4}\cmidrule(l){5-7}
+Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
+\midrule
+$\sigma$ & \Sexpr{mean(flood.zinb$sigma)} &
+  \Sexpr{median(flood.zinb$sigma)}&
+  (\Sexpr{quantile(flood.zinb$sigma, probs=c(.025))[1]}, 
+    \Sexpr{quantile(flood.zinb$sigma, probs=c(.975))[1]})&
+  \Sexpr{mean(tornado.zinb$sigma)} &
+  \Sexpr{median(tornado.zinb$sigma)}&
+  (\Sexpr{quantile(tornado.zinb$sigma, probs=c(.025))[1]}, 
+    \Sexpr{quantile(tornado.zinb$sigma, probs=c(.975))[1]})\\
+$p$ &  \Sexpr{mean(flood.zinb$p)} &
+  \Sexpr{median(flood.zinb$p)}&
+  (\Sexpr{quantile(flood.zinb$p, probs=c(.025))[1]}, 
+    \Sexpr{quantile(flood.zinb$p, probs=c(.975))[1]})&
+  \Sexpr{mean(tornado.zinb$p)} &
+  \Sexpr{median(tornado.zinb$p)}&
+  (\Sexpr{quantile(tornado.zinb$p, probs=c(.025))[1]}, 
+    \Sexpr{quantile(tornado.zinb$p, probs=c(.975))[1]})\\
+$r$ &  \Sexpr{mean(flood.zinb$r)} &
+  \Sexpr{median(flood.zinb$r)}&
+  (\Sexpr{quantile(flood.zinb$r, probs=c(.025))[1]}, 
+    \Sexpr{quantile(flood.zinb$r, probs=c(.975))[1]})&
+  \Sexpr{mean(tornado.zinb$r)} &
+  \Sexpr{median(tornado.zinb$r)}&
+  (\Sexpr{quantile(tornado.zinb$r, probs=c(.025))[1]}, 
+    \Sexpr{quantile(tornado.zinb$r, probs=c(.975))[1]})\\
+\bottomrule
+\end{tabular}
+\end{table}
+\begin{figure}
+	\centering
+<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=4.75, fig.width=7>>=
+B<-length(flood.zinb$sigma)
+Bt<-length(tornado.zinb$sigma)
+par(mfrow=c(4, 3))
+par(mgp=c(1.25,.5,0))
+par(mar=c(3.1,4.1,2.5,.1))
+plot(cumsum(flood.zinb$sigma)/1:B, 
+     main = expression(paste('Cum. Mean (Flood) ',sigma)),
+     ylab = 'Cumulative Mean',
+     type='l')
+plot(cumsum(flood.zinb$r)/1:B, 
+     main = expression(paste('Cum. Mean (Flood) ',r)),
+     ylab = 'Cumulative Mean',
+     type='l')
+plot(cumsum(flood.zinb$p)/1:B, 
+     type='l', 
+     main = expression(paste('Cum. Mean (Flood) ',p)),
+     ylab = 'Cumulative Mean')
+acf(flood.zinb$sigma, 
+    main = expression(paste('ACF (Flood) ',sigma)))
+acf(flood.zinb$r,
+    main = expression(paste('ACF (Flood) ',r)))
+acf(flood.zinb$p,
+    main = expression(paste('ACF (Flood) ',p)))
+plot(cumsum(tornado.zinb$sigma)/1:Bt, 
+     main = expression(paste('Cum. Mean (Tornado) ',sigma)),
+     ylab = 'Cumulative Mean',
+     type='l')
+plot(cumsum(tornado.zinb$r)/1:Bt, 
+     main = expression(paste('Cum. Mean (Tornado) ',r)),
+     ylab = 'Cumulative Mean',
+     type='l')
+plot(cumsum(tornado.zinb$p)/1:Bt, 
+     type='l', 
+     main = expression(paste('Cum. Mean (Tornado) ',p)),
+     ylab = 'Cumulative Mean')
+acf(tornado.zinb$sigma, 
+    main = expression(paste('ACF (Tornado) ',sigma)))
+acf(tornado.zinb$r,
+    main = expression(paste('ACF (Tornado) ',r)))
+acf(tornado.zinb$p,
+    main = expression(paste('ACF (Tornado) ',p)))
+@
+	\caption{Autocorrelation plots, index plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood and the tornado data.}
+	\label{f:zinbconvergence}
+\end{figure}
+
+\begin{figure}
+	\centering
+<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3>>=
+par(mfrow=c(1,3))
+double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
+                    expression(paste(sigma, ' Posterior Density ')), 
+                    expression(sigma), 
+                    expression(p*group("(",paste(sigma, "|", X),")")),
+                    col=c('blue', 'black'))
+double.density.plot(flood.zinb$p, tornado.zinb$p,
+                    expression(paste(p, ' Posterior Density ')), 
+                    expression(p), 
+                    expression(p*group("(",paste(p, "|", X),")")),
+                    col=c('blue', 'black'))
+double.density.plot(flood.zinb$r, tornado.zinb$r,
+                    expression(paste(r, ' Posterior Density ')), 
+                    expression(r), 
+                    expression(p*group("(",paste(r, "|", X),")")),
+                    col=c('blue', 'black'))
+@
+	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
+	\label{f:zinbdensity}
+\end{figure} 
 \section{Discussion}
@@
 This appendix includes the code used to implement the models.
+\subsection{Zero Inflated Negative Binomial code}
+\label{a:cZINB}
+<<zinbsampler-source, echo=TRUE, ref.label='zinbsampler'>>=
+@
 
@@
 \end{align*}
-
 \end{document}

### 2017-05-10 paper.tex
finish zing results

--- 2017-05-09 paper.tex
@@
 \setlength{\marginparwidth}{2cm}
-\usepackage{Sweave}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-\input{paper-concordance}
+
 \begin{center}
@@
 
+The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution.
+For the source code of the sampler, please see \ref{a:cZINB}.
+
+This model suffered severely from the curse of dimensionality.
+It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary.
+Autocorrelation and running mean plots for both the Tornado and the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+
+In the end $650000$ samples were drawn originally for each model, with a burn in of $20000$ and one of every $70$ samples was retained, resulting in a final sample size of $9000$.
+The flood model had an acceptance rate of $0.193$ and the tornado model had an acceptance rate of $0.193$.
+The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
+
+The DIC for floods and tornadoes respectively were $10804$ and $7166$. I was unable to calculate WIACs because it was too computationally expensive.
+\begin{table}
+\centering
+\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
+\label{t:ZINBparams}
+\begin{tabular}{lcccccc}
+\toprule
+&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
+\cmidrule(r){2-4}\cmidrule(l){5-7}
+Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
+\midrule
+$\sigma$ & 0.305 &
+  0.285&
+  (0.014, 
+    0.712)&
+  0.29 &
+  0.267&
+  (0.012, 
+    0.682)\\
+$p$ &  0.555 &
+  0.555&
+  (0.519, 
+    0.591)&
+  0.862 &
+  0.863&
+  (0.841, 
+    0.882)\\
+$r$ &  0.023 &
+  0.02&
+  (0.014, 
+    0.052)&
+  0.015 &
+  0.013&
+  (0.009, 
+    0.03)\\
+\bottomrule
+\end{tabular}
+\end{table}
+\begin{figure}
+	\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/floodconvergence-1}
+
+\end{knitrout}
+	\caption{Autocorrelation plots, index plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood and the tornado data.}
+	\label{f:zinbconvergence}
+\end{figure}
+
+\begin{figure}
+	\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/zinbdensity-1}
+
+\end{knitrout}
+	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
+	\label{f:zinbdensity}
+\end{figure} 
 \section{Discussion}
@@
 This appendix includes the code used to implement the models.
+\subsection{Zero Inflated Negative Binomial code}
+\label{a:cZINB}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{logpost} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{sigma}\hlstd{,} \hlkwc{p}\hlstd{,} \hlkwc{r}\hlstd{,} \hlkwc{Z}\hlstd{,} \hlkwc{N}\hlstd{,} \hlkwc{S}\hlstd{,} \hlkwc{nonzero}\hlstd{)\{}
+  \hlstd{lp} \hlkwb{<-} \hlstd{Z}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{/}\hlstd{sigma}\hlopt{-}\hlnum{1}\hlstd{)}\hlopt{*}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}\hlopt{^}\hlstd{r)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{Z}\hlopt{*}\hlkwd{log}\hlstd{(sigma)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{sigma)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlstd{r}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{S}\hlopt{*}\hlkwd{log}\hlstd{(p)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{-} \hlkwd{log}\hlstd{(r)}\hlopt{/}\hlnum{2}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{-} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlkwd{lgamma}\hlstd{(r)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlkwd{sum}\hlstd{(}\hlkwd{lgamma}\hlstd{(r}\hlopt{+}\hlstd{nonzero))}
+  \hlkwd{return}\hlstd{(lp)}
+\hlstd{\}}
+
+\hlstd{loglik} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{sigma}\hlstd{,} \hlkwc{p}\hlstd{,} \hlkwc{r}\hlstd{,} \hlkwc{Z}\hlstd{,} \hlkwc{N}\hlstd{,} \hlkwc{S}\hlstd{,} \hlkwc{nonzero}\hlstd{)\{}
+  \hlstd{lp} \hlkwb{<-} \hlstd{Z}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{/}\hlstd{sigma}\hlopt{-}\hlnum{1}\hlstd{)}\hlopt{*}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}\hlopt{^}\hlstd{r)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{Z}\hlopt{*}\hlkwd{log}\hlstd{(sigma)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{sigma)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlstd{r}\hlopt{*}\hlkwd{log}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlstd{S}\hlopt{*}\hlkwd{log}\hlstd{(p)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{-} \hlkwd{sum}\hlstd{(}\hlkwd{lgamma}\hlstd{(nonzero}\hlopt{+}\hlnum{1}\hlstd{))}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{-} \hlstd{(N}\hlopt{-}\hlstd{Z)}\hlopt{*}\hlkwd{lgamma}\hlstd{(r)}
+  \hlstd{lp} \hlkwb{<-} \hlstd{lp} \hlopt{+} \hlkwd{sum}\hlstd{(}\hlkwd{lgamma}\hlstd{(r}\hlopt{+}\hlstd{nonzero))}
+  \hlkwd{return}\hlstd{(lp)}
+\hlstd{\}}
+
+\hlstd{g} \hlkwb{<-}\hlkwa{function}\hlstd{(}\hlkwc{sigma}\hlstd{,} \hlkwc{p}\hlstd{,} \hlkwc{r}\hlstd{,}
+             \hlkwc{sigma.mean}\hlstd{,} \hlkwc{p.mean}\hlstd{,} \hlkwc{r.mean}\hlstd{,}
+             \hlkwc{sigma.var}\hlstd{,} \hlkwc{p.var}\hlstd{,} \hlkwc{r.var}\hlstd{)\{}
+    \hlkwd{return}\hlstd{(}\hlkwd{dtruncnorm}\hlstd{(sigma,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, sigma.mean,} \hlkwd{sqrt}\hlstd{(sigma.var))}\hlopt{*}
+           \hlkwd{dtruncnorm}\hlstd{(p,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, p.mean,} \hlkwd{sqrt}\hlstd{(p.var))}\hlopt{*}
+           \hlkwd{dtruncnorm}\hlstd{(r,} \hlnum{0}\hlstd{,} \hlnum{Inf}\hlstd{, r.mean,} \hlkwd{sqrt}\hlstd{(r.var)))}
+\hlstd{\}}
+
+\hlstd{zinb.sampler} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{df}\hlstd{,} \hlkwc{event_type}\hlstd{,} \hlkwc{chainlen}\hlstd{,}
+                         \hlkwc{r.var}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{p.var}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{sigma.var}\hlstd{=}\hlnum{1}\hlstd{,}
+                         \hlkwc{burnin}\hlstd{=}\hlnum{0}\hlstd{,} \hlkwc{thinning}\hlstd{=}\hlnum{1}\hlstd{)\{}
+  \hlstd{X} \hlkwb{<-} \hlstd{df[df}\hlopt{$}\hlstd{EVENT_TYPE}\hlopt{==}\hlstd{event_type,}\hlstr{"DEATHS_DIRECT"}\hlstd{]}
+  \hlstd{N} \hlkwb{<-} \hlkwd{length}\hlstd{(X)}
+  \hlstd{Z} \hlkwb{<-} \hlkwd{sum}\hlstd{(X}\hlopt{==}\hlnum{0}\hlstd{)}
+  \hlstd{S} \hlkwb{<-} \hlkwd{sum}\hlstd{(X)}
+  \hlstd{nonzero} \hlkwb{<-} \hlstd{X[X}\hlopt{!=}\hlnum{0}\hlstd{]}
+  \hlstd{B} \hlkwb{<-} \hlstd{chainlen}
+  \hlstd{b} \hlkwb{<-} \hlstd{burnin}
+  \hlstd{r.sd} \hlkwb{<-} \hlkwd{sqrt}\hlstd{(r.var)}
+  \hlstd{p.sd} \hlkwb{<-} \hlkwd{sqrt}\hlstd{(p.var)}
+  \hlstd{sigma.sd} \hlkwb{<-} \hlkwd{sqrt}\hlstd{(sigma.var)}
+
+  \hlstd{r.array} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+  \hlstd{sigma.array} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+  \hlstd{p.array} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+  \hlstd{ar.array} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+  \hlstd{posts} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+
+  \hlstd{r.array[}\hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{.5}
+  \hlstd{p.array[}\hlnum{1}\hlstd{]} \hlkwb{<-} \hlkwd{min}\hlstd{(}\hlkwd{mean}\hlstd{(X),} \hlnum{.5}\hlstd{)}
+  \hlstd{sigma.array[}\hlnum{1}\hlstd{]} \hlkwb{<-} \hlkwd{mean}\hlstd{(X}\hlopt{==}\hlnum{0}\hlstd{)}\hlopt{/}\hlnum{2}
+  \hlcom{#print(c(r.array[1], p.array[1], sigma.array[1]))}
+  \hlstd{posts[}\hlnum{1}\hlstd{]} \hlkwb{<-} \hlkwd{logpost}\hlstd{(sigma.array[}\hlnum{1}\hlstd{], p.array[}\hlnum{1}\hlstd{], r.array[}\hlnum{1}\hlstd{], Z, N, S, nonzero)}
+  \hlkwa{for}\hlstd{(i} \hlkwa{in} \hlnum{2}\hlopt{:}\hlstd{chainlen)\{}
+    \hlstd{sigma.star} \hlkwb{<-} \hlkwd{rtruncnorm}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, sigma.array[i}\hlopt{-}\hlnum{1}\hlstd{], sigma.sd)}
+    \hlstd{p.star} \hlkwb{<-} \hlkwd{rtruncnorm}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, p.array[i}\hlopt{-}\hlnum{1}\hlstd{], p.sd)}
+    \hlstd{r.star} \hlkwb{<-} \hlkwd{rtruncnorm}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{Inf}\hlstd{, r.array[i}\hlopt{-}\hlnum{1}\hlstd{], r.sd)}
+    \hlstd{lpost} \hlkwb{<-}\hlkwd{logpost}\hlstd{(sigma.star, p.star, r.star, Z, N, S, nonzero)}
+    \hlstd{reject.prob} \hlkwb{<-} \hlkwd{exp}\hlstd{(lpost}\hlopt{-}\hlstd{posts[i}\hlopt{-}\hlnum{1}\hlstd{])}\hlopt{*}
+                   \hlkwd{dtruncnorm}\hlstd{(sigma.array[i}\hlopt{-}\hlnum{1}\hlstd{],} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{,}
+                              \hlstd{sigma.star, sigma.sd)}\hlopt{*}
+                   \hlkwd{dtruncnorm}\hlstd{(p.array[i}\hlopt{-}\hlnum{1}\hlstd{],} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, p.star, p.sd)}\hlopt{*}
+                   \hlkwd{dtruncnorm}\hlstd{(r.array[i}\hlopt{-}\hlnum{1}\hlstd{],} \hlnum{0}\hlstd{,} \hlnum{Inf}\hlstd{, r.star, r.sd)}\hlopt{/}
+                  \hlstd{(}\hlkwd{dtruncnorm}\hlstd{(sigma.star,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{,}
+                              \hlstd{sigma.array[i}\hlopt{-}\hlnum{1}\hlstd{], sigma.sd)}\hlopt{*}
+                     \hlkwd{dtruncnorm}\hlstd{(p.star,} \hlnum{0}\hlstd{,} \hlnum{1}\hlstd{, p.array[i}\hlopt{-}\hlnum{1}\hlstd{], p.sd)}\hlopt{*}
+                     \hlkwd{dtruncnorm}\hlstd{(r.star,} \hlnum{0}\hlstd{,} \hlnum{Inf}\hlstd{, r.array[i}\hlopt{-}\hlnum{1}\hlstd{], r.sd))}
+    \hlstd{u} \hlkwb{<-} \hlkwd{runif}\hlstd{(}\hlnum{1}\hlstd{,}\hlnum{0}\hlstd{,}\hlnum{1}\hlstd{)}
+    \hlkwa{if}\hlstd{(u} \hlopt{<} \hlkwd{min}\hlstd{(reject.prob,} \hlnum{1}\hlstd{))\{}
+      \hlstd{r.array[i]} \hlkwb{<-} \hlstd{r.star}
+      \hlstd{p.array[i]} \hlkwb{<-} \hlstd{p.star}
+      \hlstd{sigma.array[i]} \hlkwb{<-} \hlstd{sigma.star}
+      \hlstd{ar.array[i]} \hlkwb{<-} \hlnum{1}
+      \hlstd{posts[i]} \hlkwb{<-} \hlstd{lpost}
+    \hlstd{\}}
+    \hlkwa{else}\hlstd{\{}
+      \hlstd{r.array[i]} \hlkwb{<-} \hlstd{r.array[i}\hlopt{-}\hlnum{1}\hlstd{]}
+      \hlstd{p.array[i]} \hlkwb{<-} \hlstd{p.array[i}\hlopt{-}\hlnum{1}\hlstd{]}
+      \hlstd{sigma.array[i]} \hlkwb{<-} \hlstd{sigma.array[i}\hlopt{-}\hlnum{1}\hlstd{]}
+      \hlstd{posts[i]} \hlkwb{<-} \hlstd{posts[i}\hlopt{-}\hlnum{1}\hlstd{]}
+    \hlstd{\}}
+  \hlstd{\}}
+  \hlstd{inds} \hlkwb{<-} \hlkwd{seq}\hlstd{(b}\hlopt{+}\hlnum{1}\hlstd{, B,} \hlkwc{by}\hlstd{=thinning)}
+  \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(}\hlkwc{sigma}\hlstd{=sigma.array[inds],}
+              \hlkwc{p}\hlstd{=p.array[inds],}
+              \hlkwc{r}\hlstd{=r.array[inds],}
+              \hlkwc{ar}\hlstd{=ar.array))}
+\hlstd{\}}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 

### 2017-05-10 zinb_sampler.R

@@ -1,3 +1,4 @@
+# ---- zinbsampler ----
 logpost <- function(sigma, p, r, Z, N, S, nonzero){
   lp <- Z*log(1 + (1/sigma-1)*(1-p)^r)
   lp <- lp + Z*log(sigma)
@@ -10,13 +11,16 @@ logpost <- function(sigma, p, r, Z, N, S, nonzero){
   return(lp)
 }
 
-sample <- function(r.mean, p.mean, sigma.mean, r.var, p.var, sigma.var){
-  sigma.star <- rtruncnorm(1, 0, 1, sigma.mean, sqrt(sigma.var))
-  #rbeta(1, sigma.mean*sigma.N, (1-sigma.mean)*sigma.N)
-  p.star <- rtruncnorm(1, 0, 1, p.mean, sqrt(p.var))
-  #rbeta(1, p.mean*p.N, (1-p.mean)*p.N)
-  r.star <- rtruncnorm(1, 0, Inf, r.mean, sqrt(r.var))
-  return(c(sigma.star, p.star, r.star))
+loglik <- function(sigma, p, r, Z, N, S, nonzero){
+  lp <- Z*log(1 + (1/sigma-1)*(1-p)^r)
+  lp <- lp + Z*log(sigma)
+  lp <- lp + (N-Z)*log(1-sigma)
+  lp <- lp + (N-Z)*r*log(1-p)
+  lp <- lp + S*log(p)
+  lp <- lp - sum(lgamma(nonzero+1))
+  lp <- lp - (N-Z)*lgamma(r)
+  lp <- lp + sum(lgamma(r+nonzero))
+  return(lp)
 }
 
 g <-function(sigma, p, r, 
@@ -28,107 +32,134 @@ g <-function(sigma, p, r,
 }
 
 
-zinb.sampler <- function(df, year, event_type, chainlen, 
+zinb.sampler <- function(df, event_type, chainlen, 
                          r.var=1, p.var=1, sigma.var=1, 
-                         burnin=0, thinning=1, nchains=1, random=TRUE){
-  X <- df[df$YEAR>year&df$EVENT_TYPE==event_type,"DEATHS_DIRECT"]
+                         burnin=0, thinning=1){
+  X <- df[df$EVENT_TYPE==event_type,"DEATHS_DIRECT"]
   N <- length(X)
   Z <- sum(X==0)
   S <- sum(X)
   nonzero <- X[X!=0]
[…]

### 2017-05-10 paper.Rnw

@@
 \noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+
+\[
+\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
+\]
+
+\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
+
+\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
+
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
+
+\noindent Neither of these distributions is recognizable.
+We can use a Metropolis-Hastings algorithm to sample from both of them.
+We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
+We will tune them to obtain a better acceptance rate.
+
+The Zero Inflated Poisson (ZIP) model has two parameters.
+The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
+For a single observation x, the probability density is:
+
+\[
+p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
+\]
+
+\noindent We can write the likelihood as follows:
+$$
+L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
+$$
+
+\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
 
@@
 \label{ss:rZiPoisson}
+
+In tuning the parameters of the two models, slightly different values were chosen for the two different event types.
+The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods.
+The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model.
+
+For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are as follows:
+
+\[\]
+\noindent Table 1: Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types.
+The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
+The third does the same for p.
+All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.
+
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$ & p  \\ \hline
+    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
+    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
+    \end{tabular}
+\end{center}
 

### 2017-05-10 paper.Rnw
Add table for DIC/WAIC some minor clean ups

@@
 
-\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
-
-\[
-\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]
-\]
-
-\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
-
-\[
-p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
-\]
-
-\[
-p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
-\]
-
-\noindent Neither of these distributions is recognizable.
-We can use a Metropolis-Hastings algorithm to sample from both of them.
-We will use a beta distribution as a proposal for p, and a gamma for $\lambda$.
-We will tune them to obtain a better acceptance rate.
-
-The Zero Inflated Poisson (ZIP) model has two parameters.
-The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model.
-For a single observation x, the probability density is:
-
-\[
-p(x|p, \lambda) = pI_{x=0}(x) + (1-p)\frac{e^{-\lambda}\lambda^x}{x!}
-\]
-
-\noindent We can write the likelihood as follows:
-$$
-L(p, \lambda|X) = \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]
-$$
-
 \noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
@@
 
-\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
-
-\[
-p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
-\]
-
-\[
-p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
-\]
-
-\noindent Neither of these distributions is recognizable.
+\noindent Neither of the full conditional distributions is recognizable (see \ref{a:dZIP}).
 We can use a Metropolis-Hastings algorithm to sample from both of them.
@@
 
-For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are as follows:
+For each variable of interest, 20,000 samples were taken.
+As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in.
+The results are summarized in Table~\ref{t:rZIP}.
 
-\[\]
-\noindent Table 1: Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types.
+\begin{table}
+    \centering
+    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types.
 The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
 The third does the same for p.
-All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.
-
-\begin{center}
+All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
+    \label{t:rZIP}
     \begin{tabular}{| l | l | l |}
@@
     \end{tabular}
-\end{center}
+\end{table}
 
@@
 <<zinbparams, echo=FALSE, cache=TRUE>>=
-samplesize <- 650000
-burnin <- 20000
-thin <- 70
+samplesize <- 1030000
+burnin <- 30000
+thin <- 100
 r.var=.00002
@@
 It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary.
-Autocorrelation and running mean plots for both the Tornado and the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+Autocorrelation and running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
 <<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
@@
 	\centering
-<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=4.75, fig.width=7>>=
+<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3.5>>=
 B<-length(flood.zinb$sigma)
 Bt<-length(tornado.zinb$sigma)
-par(mfrow=c(4, 3))
+par(mfrow=c(2, 3))
 par(mgp=c(1.25,.5,0))
@@
     main = expression(paste('ACF (Flood) ',p)))
-plot(cumsum(tornado.zinb$sigma)/1:Bt, 
-     main = expression(paste('Cum. Mean (Tornado) ',sigma)),
-     ylab = 'Cumulative Mean',
-     type='l')
-plot(cumsum(tornado.zinb$r)/1:Bt, 
-     main = expression(paste('Cum. Mean (Tornado) ',r)),
-     ylab = 'Cumulative Mean',
-     type='l')
-plot(cumsum(tornado.zinb$p)/1:Bt, 
-     type='l', 
-     main = expression(paste('Cum. Mean (Tornado) ',p)),
-     ylab = 'Cumulative Mean')
-acf(tornado.zinb$sigma, 
-    main = expression(paste('ACF (Tornado) ',sigma)))
-acf(tornado.zinb$r,
-    main = expression(paste('ACF (Tornado) ',r)))
-acf(tornado.zinb$p,
-    main = expression(paste('ACF (Tornado) ',p)))
+# plot(cumsum(tornado.zinb$sigma)/1:Bt, 
+#      main = expression(paste('Cum. Mean (Tornado) ',sigma)),
+#      ylab = 'Cumulative Mean',
+#      type='l')
+# plot(cumsum(tornado.zinb$r)/1:Bt, 
+#      main = expression(paste('Cum. Mean (Tornado) ',r)),
+#      ylab = 'Cumulative Mean',
+#      type='l')
+# plot(cumsum(tornado.zinb$p)/1:Bt, 
+#      type='l', 
+#      main = expression(paste('Cum. Mean (Tornado) ',p)),
+#      ylab = 'Cumulative Mean')
+# acf(tornado.zinb$sigma, 
+#     main = expression(paste('ACF (Tornado) ',sigma)))
+# acf(tornado.zinb$r,
+#     main = expression(paste('ACF (Tornado) ',r)))
+# acf(tornado.zinb$p,
+#     main = expression(paste('ACF (Tornado) ',p)))
 @
-	\caption{Autocorrelation plots, index plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood and the tornado data.}
+	\caption{Autocorrelation plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data.
+The tornado plots are not pictured, but are similar in character.}
 	\label{f:zinbconvergence}
@@
 \label{s:discussion}
+The goal of this analysis was to determine which of the four models was the best fit for the data.
+In order to make this determination the DIC and WAIC was calculated for each model on both tornados and flash floods.
+The results are summarized in Table~\ref{t:evalresults}.
+
+\begin{table}
+\centering
+\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data.
+In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
+\label{t:evalresults}
+\begin{tabular}{lcccc}
+\toprule
+&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
+\cmidrule(r){2-3}\cmidrule(l){4-5}
+Model & DIC &WAIC & DIC &WAIC\\
+\midrule
+Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
+Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
+ZIP & ZIP FLOOD DIC& ZIP FLOOD WAIC &ZIP TORNADO DIC&ZIP TORNADO WAIC\\
+ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 This appendix will include details on the calculations required to derive our models.
+\subsection{Zero Inflated Poisson Derivation}
+\label{a:dZIP}
+\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:
+
+\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
+
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
 \subsection{Zero Inflated Negative Binomial Derivation}

### 2017-05-10 presentation.tex
Add latex template for presentation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> {

	% The Beamer class comes with a number of default slide themes
	% which change the colors and layouts of slides. Below this is a list
	% of all the themes, uncomment each in turn to see what they look like.

	\usetheme{default}
	%\usetheme{AnnArbor}
	%\usetheme{Antibes}
	%\usetheme{Bergen}
	%\usetheme{Berkeley}
	%\usetheme{Berlin}
	%\usetheme{Boadilla}
	%\usetheme{CambridgeUS}
	%\usetheme{Copenhagen}
	%\usetheme{Darmstadt}
	%\usetheme{Dresden}
	%\usetheme{Frankfurt}
	%\usetheme{Goettingen}
	%\usetheme{Hannover}
	%\usetheme{Ilmenau}
	%\usetheme{JuanLesPins}
	%\usetheme{Luebeck}
	%\usetheme{Madrid}
	%\usetheme{Malmoe}
	%\usetheme{Marburg}
	%\usetheme{Montpellier}
	%\usetheme{PaloAlto}
	%\usetheme{Pittsburgh}
	%\usetheme{Rochester}
	%\usetheme{Singapore}
	%\usetheme{Szeged}
	%\usetheme{Warsaw}

	% As well as themes, the Beamer class has a number of color themes
	% for any slide theme. Uncomment each of these in turn to see how it
	% changes the colors of your current slide theme.

	%\usecolortheme{albatross}
	\usecolortheme{beaver}
	%\usecolortheme{beetle}
	%\usecolortheme{crane}
	%\usecolortheme{dolphin}
	%\usecolortheme{dove}
	%\usecolortheme{fly}
	%\usecolortheme{lily}
	%\usecolortheme{orchid}
	%\usecolortheme{rose}
	%\usecolortheme{seagull}
	%\usecolortheme{seahorse}
	%\usecolortheme{whale}
	%\usecolortheme{wolverine}

	%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
	%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

	%\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{sidecap}%[capbesideposition=inside, facing=yes,capbesidesep=quad]
\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Four Models]{Four Models for Severe Weather Deaths} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Jason Michaels (jam521), Niko Paulson (ndp32), \\Miranda Seitz-McLeese (mgs85) } % Your name
%\institute[UCLA] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
%{
%	University of California \\ % Your institution for the title page
%	\medskip
%	\textit{john@smith.com} % Your email address
%}
\date{May 11, 2017} % Date, can be changed to a custom date

\begin{document}

	\begin{frame}
		\titlepage % Print the title page as the first slide
	\end{frame}

	\begin{frame}
		\frametitle{Overview} % Table of contents slide, comment this block out to remove it
		\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
	\end{frame}

	%----------------------------------------------------------------------------------------
	%	PRESENTATION SLIDES
	%----------------------------------------------------------------------------------------
	\section{Introduction} 
	\begin{frame}
		\frametitle{Introduction: Goals and Motivation}
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Data}
		The data for this project was taken from the NOAA severe weather events dataset. 
		In its original form this dataset includes a variety of measurements for all forms of severe weather in the United States dating from 1950. In the interest of time we restricted the dataset to:
		\begin{itemize}
			\item Deaths directly attributable to Tornadoes and Flash floods 
			\item Occurred between 1996-2016
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Base Models}
		% description of negative binomials and poisson models
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Zero Inflated models}
		% overview of what a zero inflated models are 
	\end{frame}
	\begin{frame}
		\frametitle{Introduction: Four Models}
		% list the four models
	\end{frame}

	\section{Negative Binomial Model}
	\begin{frame}
		\frametitle{Negative Binomial}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Negative Binomial}
		% Slide two
	\end{frame}

	\section{Poisson Model}
	\begin{frame}
		\frametitle{Poisson Model}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Poisson Model}
		% Slide two
	\end{frame}

	\section{Zero Inflated Poisson Model}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		% Slide two
	\end{frame}

	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		% Slide one
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		% Slide two
	\end{frame}

	\section{Conclusions}
	\begin{frame}
		\frametitle{Conclusions: Model Evaluation Table}
		% Waic dic table goes here
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions: Interpreting the Best Model}
		% Waic dic table goes here
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions: Future Work}
		% List of things we might have done if we'd had time
		\begin{enumerate}
			\item{Re parameterize the ZINB model/look into other techniques.}
			\item{See if the results hold on other extreme weather events.}
		\end{enumerate}
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions}
		Questions?
	\end{frame}

\end{document}

### 2017-05-10 paper.Rnw
Fill in ZINB slides tweak intro paper

@@
 The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
-For this project we focused on the deaths directly attributable to the event.
+For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
 Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
@@
 For a single $X$ the probability density is: 
-$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}.$$
+$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r.$$
 We take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. For a full derivation, see \ref{a:dZINB}. My posterior is:

### 2017-05-10 presentation.tex
Fill in ZINB slides tweak intro paper

@@
 		\frametitle{Introduction: Goals and Motivation}
+		Understanding how and at what rate severe weather events become lethal in the United States has tremendous health impacts.
+Understanding the distribution of death counts can help meteorologists better predict outcomes.
 	\end{frame}
@@
 	\begin{frame}
-		\frametitle{Introduction: Base Models}
-		% description of negative binomials and poisson models
-	\end{frame}
-	\begin{frame}
-		\frametitle{Introduction: Zero Inflated models}
-		% overview of what a zero inflated models are 
+		\frametitle{Introduction: Zero Inflated Models}
+			Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants.
+A zero inflated model is a combination of the original distribution and a point mass at zero.
+If the original distribution is $p(x)$ then $$z(x) = \begin{cases}
+			\sigma + (1-\sigma)*p(0) &x=0\\
+			(1-\sigma)*p(x) &x\ne0
+			\end{cases}$$
 	\end{frame}
@@
 		\frametitle{Introduction: Four Models}
-		% list the four models
+				In this paper we compare four possible models for the deaths: 
+				\begin{enumerate}
+					\item Poisson
+					\item Negative Binomial
+					\item Zero Inflated Poisson
+					\item Zero Inflated Negative Binomial
+				\end{enumerate}
 	\end{frame}
@@
 		\frametitle{Zero Inflated Poisson Model}
-		% Slide one
+
 	\end{frame}
@@
 		\frametitle{Zero Inflated Poisson Model}
-		% Slide two
+
 	\end{frame}
@@
 		\frametitle{Zero Inflated Negative Binomial Model}
-		% Slide one
+		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
+		For a single $X$ the probability density is: 
+		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
 	\end{frame}
@@
 		\frametitle{Zero Inflated Negative Binomial Model}
-		% Slide two
+		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
+		This gave a posterior of:
+		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
+		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Negative Binomial Model:Results}
+		\begin{itemize}
+			\item Curse of Dimensionality; very slow convergence.
+			\item High autocorrelation
+			\item Low confidence in results
+		\end{itemize}
+		% parameter table goes here
 	\end{frame}
@@
 		\begin{enumerate}
-			\item{Re parameterize the ZINB model/look into other techniques.}
+			\item{Re-parameterize the ZINB model/look into other techniques to improve convergence.}
 			\item{See if the results hold on other extreme weather events.}

### 2017-05-10 paper.Rnw
Updated Methods, Results, and DIC for ZIP.

@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 <<imports, cache=FALSE, echo=FALSE>>= 
@@
     Event Type & $\lambda$ & p  \\ \hline
-    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
-    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
+    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
+    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
     \end{tabular}
 \end{table}
+
+\includegraphics{figure/ZIP_Flood_Conv.pdf}
+
+\noindent Figure 1. Cumulative mean plots for the posterior samples of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.
+
+\includegraphics{figure/ZIP_Density.pdf}
+
+\noindent Figure 2. Posterior density plots for flood parameters (in blue), and tornado parameters (black).
 
@@
 Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
-ZIP & ZIP FLOOD DIC& ZIP FLOOD WAIC &ZIP TORNADO DIC&ZIP TORNADO WAIC\\
+ZIP & 3845.915& ZIP FLOOD WAIC &4759.998 &ZIP TORNADO WAIC\\
 ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\

### 2017-05-10 paper.Rnw
Added ZIP Code

@@
 @
+\subsection{Zero Inflated Poisson code}
+<<eval = FALSE>>=
+library(dplyr)
+tornado <- data %>% filter(EVENT_TYPE == "Tornado") %>% select(DEATHS_DIRECT)
+flood <- data %>% filter(EVENT_TYPE == "Flash Flood") %>% select(DEATHS_DIRECT)
+
+# Precompute values to use in posterior function
+zeros <- tornado %>% filter(DEATHS_DIRECT == 0) %>% select(DEATHS_DIRECT)
+ones <- tornado %>% filter(DEATHS_DIRECT != 0) %>% select(DEATHS_DIRECT)
+n <- nrow(ones)
+t0_exp <- (zeros)/factorial(zeros)
+l_t1_exp <- ones-1/2
+
+# Define Posterior Functions
+log_lam <- function(y, p, lambda){
+  t0 <- sum(log(p + (1-p)*exp(-1*lambda)*(lambda+2e-10)^t0_exp))
+  t1 <- sum(log(exp(-1*lambda)*lambda^l_t1_exp))
+  return(t0+t1)
+}
+
+log_p <- function(y, p, lambda){
+  t0 <- sum(log(p + (1-p)*exp(-1*lambda)*(lambda+2e-10)^t0_exp))
+  t1_vec <- rep(1-p, n)
+  t1 <- sum(log(t1_vec))
+  return(t0+t1)
+}
+
+### Sampler for Tornado
+set.seed(2011)
+
+l_last <- 1
+p_last <- .98
+
+B <- 20000
+lambdas <- rep(0, B)
+ps <- rep(0, B)
+
+l_accept <- rep(FALSE, B)
+p_accept <- rep(FALSE, B)
+
+for(t in 1:B){
+  lambda <- rgamma(1, 3, 2)
+  p <- rbeta(1, 1940, 60)
+  ro_lam <- log_lam(tornado, p_last, lambda) - log(dgamma(lambda, 3, 2)) -
+    log_lam(tornado, p_last, l_last) + log(dgamma(l_last, 3, 2))
+  ro_p <- log_p(tornado, p, l_last) - log(dbeta(p, 1940, 60)) -
+    log_p(tornado, p_last, l_last) + log(dbeta(p_last, 1940, 60))
+  U_lambda <- runif(1)
+  U_p <- runif(1)
+  if(log(U_lambda) < min(0, ro_lam)){
+    print(TRUE)
+    lambdas[t] <- lambda
+    lambda_last <- lambda
+    l_accept[t] <- TRUE
+  } else{
+    print(FALSE)
+    lambdas[t] <- lambda_last
+  }
+
+  if(log(U_p) < min(0, ro_p)){
+    print(TRUE)
+    ps[t] <- p
+    p_last <- p 
+    p_accept[t] <- TRUE
+  } else{
+    print(FALSE)
+    ps[t] <- p_last
+  }
+}
+
+# Do the same for flash floods
+
+# Update values for likelihood function
+zeros <- flood %>% filter(DEATHS_DIRECT == 0) %>% select(DEATHS_DIRECT)
+ones <- flood %>% filter(DEATHS_DIRECT != 0) %>% select(DEATHS_DIRECT)
+n <- nrow(ones)
+t0_exp <- (zeros)/factorial(zeros)
+l_t1_exp <- ones-1/2
+
+# Sampler:
+set.seed(2012)
+l_last <- 1
+p_last <- .98
+B <- 20000
+lambdas <- rep(0, B)
+ps <- rep(0, B)
+l_accept <- rep(FALSE, B)
+p_accept <- rep(FALSE, B)
+for(t in 1:B){
+  print(paste0("t = ", as.character(t)))
+  lambda <- rgamma(1, 1, 2)
+  print(paste0("lambda = ", as.character(lambda)))
+  p <- rbeta(1, 2945, 55)
+  print(paste0("p = ", as.character(p)))
+  ro_lam <- log_lam(flood, p_last, lambda) - log(dgamma(lambda, 1, 2)) -
+    log_lam(flood, p_last, l_last) + log(dgamma(l_last, 1, 2))
+  ro_p <- log_p(flood, p, l_last) - log(dbeta(p, 2945, 55)) -
+    log_p(flood, p_last, l_last) + log(dbeta(p_last, 2945, 55))
+  U_lambda <- runif(1)
+  U_p <- runif(1)
+  if(log(U_lambda) < min(0, ro_lam)){
+    print(TRUE)
+    lambdas[t] <- lambda
+    lambda_last <- lambda
+    l_accept[t] <- TRUE
+  } else{
+    print(FALSE)
+    lambdas[t] <- lambda_last
+  }
+
+  if(log(U_p) < min(0, ro_p)){
+    print(TRUE)
+    ps[t] <- p
+    p_last <- p 
+    p_accept[t] <- TRUE
+  } else{
+    print(FALSE)
+    ps[t] <- p_last
+  }
+}
+@
 

### 2017-05-10 paper.Rnw
Fixed Screwy ZIP DIC Code

@@
 Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
-ZIP & 3845.915& ZIP FLOOD WAIC &4759.998 &ZIP TORNADO WAIC\\
+ZIP & 10454.860& ZIP FLOOD WAIC &8727.647 &ZIP TORNADO WAIC\\
 ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\

### 2017-05-11 paper.Rnw
Edited for length, fixed problem with figures

--- 2017-05-10 paper.Rnw
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 <<imports, cache=FALSE, echo=FALSE>>= 
@@
 These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$.
-
 We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.
@@
 The results are summarized in Table~\ref{t:rZIP}.
+Convergance plots can be found in Figrue~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.
 
@@
     \centering
-    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types.
+    \caption{The posterior distributions of $\lambda$ and p, for both event types.
 The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
-The third does the same for p.
-All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
+The third does the same for p.}
     \label{t:rZIP}
@@
 
-\includegraphics{figure/ZIP_Flood_Conv.pdf}
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/ZIP_Flood_Conv.pdf}
+\caption{Cumulative mean plots for the posterior samples (ZIP model) of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.}
+\label{f:zipconv}
+\end{figure}
 
-\noindent Figure 1. Cumulative mean plots for the posterior samples of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.
-
-\includegraphics{figure/ZIP_Density.pdf}
-
-\noindent Figure 2. Posterior density plots for flood parameters (in blue), and tornado parameters (black).
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/ZIP_Density.pdf}
+\caption{Posterior density plots (ZIP model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the mean and the dashed lines indicate the 95\% confidence interval.}
+\label{f:zipdensity}
+\end{figure}
 
@@
 This model suffered severely from the curse of dimensionality.
-It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary.
-Autocorrelation and running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary.
+Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
 <<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
@@
 @
-In the end $\Sexpr{samplesize}$ samples were drawn originally for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
+In the end $\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
 The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$.
-The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
+The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.
 <<evaluation, echo=FALSE, cache=FALSE>>=
@@
 The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.
-I was unable to calculate WIACs because it was too computationally expensive.
 \begin{table}
@@
 \end{table}
-\begin{figure}
+\begin{figure}[p]
 	\centering
-<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3.5>>=
+<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=1.75>>=
 B<-length(flood.zinb$sigma)
 Bt<-length(tornado.zinb$sigma)
-par(mfrow=c(2, 3))
+par(mfrow=c(1, 3))
 par(mgp=c(1.25,.5,0))
@@
      ylab = 'Cumulative Mean')
-acf(flood.zinb$sigma, 
-    main = expression(paste('ACF (Flood) ',sigma)))
-acf(flood.zinb$r,
-    main = expression(paste('ACF (Flood) ',r)))
-acf(flood.zinb$p,
-    main = expression(paste('ACF (Flood) ',p)))
+# acf(flood.zinb$sigma, 
+#     main = expression(paste('ACF (Flood) ',sigma)))
+# acf(flood.zinb$r,
+#     main = expression(paste('ACF (Flood) ',r)))
+# acf(flood.zinb$p,
+#     main = expression(paste('ACF (Flood) ',p)))
 # plot(cumsum(tornado.zinb$sigma)/1:Bt, 
@@
 @
-	\caption{Autocorrelation plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data.
-The tornado plots are not pictured, but are similar in character.}
+	\caption{Cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
 	\label{f:zinbconvergence}
@@
 
-\begin{figure}
+\begin{figure}[p]
 	\centering
-<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3>>=
+<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=2.5, fig.width=6>>=
 par(mfrow=c(1,3))
@@
 \centering
-\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data.
-In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
+\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
 \label{t:evalresults}
-\begin{tabular}{lcccc}
+\begin{tabular}{lcc}
 \toprule
-&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
-\cmidrule(r){2-3}\cmidrule(l){4-5}
-Model & DIC &WAIC & DIC &WAIC\\
+Model & DIC Flash Flood & DIC Tornado\\
 \midrule
-Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
-Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
-ZIP & 3845.915& ZIP FLOOD WAIC &4759.998 &ZIP TORNADO WAIC\\
-ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
+Poisson & P FLOOD DIC& P TORNADO DIC\\
+Negative Binomial & NB FLOOD DIC&NB TORNADO DIC\\
+ZIP & 10454.860&8727.647 \\
+ZINB & \Sexpr{flood.dic.zinb}&\Sexpr{tornado.dic.zinb}\\
 \bottomrule
@@
 \end{table}
-
+\clearpage
 \begin{thebibliography}{2}
@@
 \end{thebibliography}
-\clearpage
+
 \appendix

### 2017-05-11 presentation.tex
added ZINB results to slides
ZINB tables

@@
 		\end{itemize}
-		% parameter table goes here
+		\begin{table}
+			\centering
+			\scalebox{0.6}{
+			\begin{tabular}{lcccccc}
+				\toprule
+				&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
+				\cmidrule(r){2-4}\cmidrule(l){5-7}
+				Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
+				\midrule
+				$\sigma$ & 0.307 &
+				0.282&
+				(0.013, 
+				0.716)&
+				0.299 &
+				0.276&
+				(0.012, 
+				0.702)\\
+				$p$ &  0.554 &
+				0.555&
+				(0.518, 
+				0.589)&
+				0.862 &
+				0.863&
+				(0.841, 
+				0.882)\\
+				$r$ &  0.024 &
+				0.02&
+				(0.014, 
+				0.052)&
+				0.015 &
+				0.013&
+				(0.009, 
+				0.033)\\
+				\bottomrule
+			\end{tabular}}
+		\end{table}
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Negative Binomial Model}
+		\begin{figure}[p]
+			\centering
+			\includegraphics[width=.9\textwidth]{figure/zinbdensity-1} 
+			\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
+		\end{figure} 
 	\end{frame}
@@
 			\item{See if the results hold on other extreme weather events.}
+			\item{Attempt to fit all data with a mixture model.}
 		\end{enumerate}

### 2017-05-11 paper_edits.Rnw
added pois/nb results and conclusion

--- 2017-05-11 paper.Rnw
@@
 This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
-The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
-For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
+We chose to use a subset of that data covering the years 1996 through 2016.
+The data coms from the NOAA website \cite{dataset}.
+For this project we focused on the deaths directly attributable to a type of weather event.
+In the interest of time, we limited our models to two types of severe weather, Tornados and Flash Floods, reponsible for the second and third most direct deaths among weather types.
 Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
@@
 This distribution does not factor nicely, so I will use the Metropolis algorithm to sample from it.
-Because this posterior does not suggest any obvious proposal distributions I will sample each independently from a normal distribution centered at $\theta^*$, and with a variance that is tuned to yeild an appropriate acceptance rate.
+Because this posterior does not suggest any obvious proposal distributions I will sample each independently from a normal distribution centered at $\theta^*$, and with a variance that is tuned to yield an appropriate acceptance rate.
 
@@
 
+Because the posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, we were able to sample from this distribution directly.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $\lambda$.
+The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $\lambda$, for both event types. The second column relays the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
+    \label{t:rPOIS}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$  \\ \hline
+    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
+    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/POIS_Density.pdf}
+\caption{Posterior density plots (Poisson model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:poisdensity}
+\end{figure}
+
 \subsection{Negative Binomial}
 \label{ss:rNBinom}
+
+As in the case of the Poisson distribution, the conditional posterior from our Negative Binomial likelihood and noninformative prior had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly.
+Meanwhile we took $r$ as known.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $p$.
+The results are summarized in Table~\ref{t:rNB}, while density plots can be found in Figure~\ref{f:nbdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $p$, for both event types. The second column relays the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
+    \label{t:rNB}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $p$  \\ \hline
+    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
+    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/NB_Density.pdf}
+\caption{Posterior density plots (Negative Binomial model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:nbdensity}
+\end{figure}
 
@@
 The results are summarized in Table~\ref{t:rZIP}.
-Convergance plots can be found in Figrue~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.
+Convergance plots can be found in Figure~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.
 
@@
 The goal of this analysis was to determine which of the four models was the best fit for the data.
-In order to make this determination the DIC and WAIC was calculated for each model on both tornados and flash floods.
+In order to make this determination the DIC was calculated for each model on both tornados and flash floods.
 The results are summarized in Table~\ref{t:evalresults}.
@@
 \midrule
-Poisson & P FLOOD DIC& P TORNADO DIC\\
-Negative Binomial & NB FLOOD DIC&NB TORNADO DIC\\
-ZIP & 10454.860&8727.647 \\
+Poisson & 14237 & 17606\\
+Negative Binomial & 101548 & 26372\\
+ZIP & 10454 & 8727 \\
 ZINB & \Sexpr{flood.dic.zinb}&\Sexpr{tornado.dic.zinb}\\
@@
 \end{table}
+
+In general, models with smaller DIC are preferred to models with larger DIC.
+From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados.
+Thus for estimating the rate of leathality of Flash Floods, Tornados, and possibly other severe weather, we should use a Zero Inflated Poisson model to fit the data.
+
 \clearpage
@@
 @
+\subsection{Poisson code}
+<<eval = FALSE>>=
+set.seed(05112017)
+
+# --- model evaluation function -----------------------------------------------------------
+DIC <- function(y, theta, loglik) {
+  theta_hat = mean(theta) #theta_hat = apply(theta, 2, mean)
+  L = loglik(y, theta_hat)
+  S = length(theta) #S = nrow(theta) #S = number of iterations
+  llSum = 0
+  for (s in 1:S) {
+    theta_s = theta[s]
+    llSum = llSum + loglik(y, theta_s)
+  }
+  P = 2 * (L - (1 / S * llSum))
+  DIC = - 2 * (L - P)
+  return(DIC)
+}
+
+# --- poisson log-likelihood ---------------------------------------------------------
+pois_ll <- function(x, lambda) {
+  ll <- sum(dpois(x, lambda, log = TRUE))
+  return(ll)
+}
+
+# --- poisson model -----------------------------------------------------------------
+pois_fit <- function(data, nsim) {
+  x <- data                            
+  xbar <- mean(x)                      
+  n <- length(x)                       
+  lambda <- rgamma(nsim, n * xbar + 1 /2, n) 
+  return(lambda)
+}
+
+# --- results - ---------------------------------------------------------------------------
+lambda_t <- pois_fit(data = TORNADO$DEATHS, nsim = 10000)
+lambda_ff <- pois_fit(data = FLASH_FLOOD$DEATHS, nsim = 10000)
+
+quantile(lambda_t, c(0.025, 0.5, 0.975))
+quantile(lambda_ff, c(0.025, 0.5, 0.975))
+
+DIC(TORNADO$DEATHS, lambda_t, pois_ll)
+DIC(FLASH_FLOOD$DEATHS, lambda_ff, pois_ll)
+@
+
+\subsection{Negative Binomial code}
+<<eval = FALSE>>=
+# --- negative binomial log-likelihood ---------------------------------------------------------
+nb_ll <- function(x, p) {
+  ll <- sum(dnbinom(x = x, prob = p, size = 1 / 4, log = TRUE))
+  return(ll)
+}
+
+# --- negative binomial model -----------------------------------------------------------------
+nb_fit <- function(data, nsim) {
+  x <- data
+  xbar <- mean(x)
+  n <- length(x)
+  r <- 1 / 4
+  p <- rbeta(nsim, n * xbar, n * r + 1 / 2) 
+  return(p)
+}
+
+# --- results - ---------------------------------------------------------------------------
+p_t <- nb_fit(data = TORNADO$DEATHS, nsim = 10000)
+p_ff <- nb_fit(data = FLASH_FLOOD$DEATHS, nsim = 10000)
+
+# --- credible intervals ----------------------------------------------------------------------
+quantile(p_t, c(0.025, 0.5, 0.975))
+quantile(p_ff, c(0.025, 0.5, 0.975))
+
+# --- dic ----------------------------------------------------------------------
+DIC(TORNADO$DEATHS, p_t, nb_ll)
+DIC(FLASH_FLOOD$DEATHS, p_ff, nb_ll)
+
+@
 

### 2017-05-11 presentation.tex
First Pass at ZIP Slide

@@
 		\frametitle{Zero Inflated Poisson Model}
-
+		The zero inflated Poisson model has two parameters p, the probability of a structural zero, and $\lambda$, the usual Poisson parameter. 
+		For a single X, the probability density is:
+		$$pI_{X=0}(X) + (1-p)\frac{\lambda^Xe^-\lambda}{X!}$$
 	\end{frame}
@@
 		\frametitle{Zero Inflated Poisson Model}
+		One way to write the likelihood is as follows:
+		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]$$
+		Bayarri, Berger, and Datta (2008)\footnotemark suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior:
+		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]$$
+		\footnotetext[1]{Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121.}
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Poisson Model}
+		We can simplify a bit to get the following conditionals:
+		\[
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+\]
 
+\[
+p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
+\]
+
+These are unrecognizable. I ended up alternatively sampling from the two using Metropolis Hastings. For lambda I settled on a gamma proposal, and for p I settled on a Beta proposal. 
 	\end{frame}
-
+	\begin{frame}
+		\frametitle{Zero Inflated Poisson Model}
+		\begin{table}
+    \centering
+    \caption{After taking 20,000 samples from each variable and discarding 10,000 as a burn-in, we obtained the following (values are the mean of the posterior sample followed by 95 percent credible interval in parentheses)}
+    \label{t:rZIP}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$ & p  \\ \hline
+    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
+    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
+    \end{tabular}
+\end{table}
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Poisson Model}
+		Convergence of parameters (flood model):
+		\centering
+		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Poisson Model}
+		Posterior Densities (Blue = Flood; Black = Tornado):
+		\centering
+		\includegraphics[width=.9\textwidth]{figure/ZIP_Density.pdf} 
+	\end{frame}	
 	\section{Zero Inflated Negative Binomial Model}

### 2017-05-11 presentation.tex
Changes to ZIP slides

@@
 		Bayarri, Berger, and Datta (2008)\footnotemark suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior:
-		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i - 1/2}}{x_i!}\bigg]$$
+		$$\prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\lambda^{-1/2}$$
 		\footnotetext[1]{Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121.}
@@
 		\[
-p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i - 1/2}\bigg]
+p(\lambda|X, p) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[e^{-\lambda}\lambda^{x_i}\bigg]\lambda^{-1/2}
 \]

### 2017-05-11 paper.Rnw
Edit for length

@@
 This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
-The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
-For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
+We chose to use a subset of that data covering the years 1996 through 2016.
+The data coms from the NOAA website \cite{dataset}.
+For this project we focused on the deaths directly attributable to a type of weather event.
+In the interest of time, we considered two types of severe weather, Tornados and Flash Floods, reponsible for the second and third most direct deaths among weather types.
 Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
 In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.
-
 The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models.
@@
 $$p(x|\lambda)=\frac{\lambda^xe^{-\lambda}}{x!}.$$
-Thus the likelihood for a sum of Poisson random variables can be written as follows:
-$$\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{X}}e^{-n\lambda}.$$ 
-Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
-$$\pi(\lambda)\propto\lambda^{1/2-1}e^{-0\cdot\lambda}.$$
-We recognize this as the kernel of an improper gamma distribution. Combining the likelihood and the prior distribtion yields the following posterior distribution:
-$$p(\lambda|X)=\lambda^{n\bar{X}+1/2-1}e^{-n\lambda}.$$
-We recognize this as the kernel of a gamma distribution, namely
+We use Jeffrey's prior, and model 
 $$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$ 
+For a full derivation see~\ref{a:dP}.
 
@@
 $$p(X|r,p)=\frac{\Gamma(r+x)}{\Gamma(r)x!}p^{x}(1-p)^{r}.$$
-
-Thus the likelihood for a sum of Negative Binomial random variables can be written as follows:
-$$\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}.$$
-Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
-$$\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}.$$
-Combining the likelihood and the prior distribtion yields the following posterior distribution:
-$$p(r,p|X)=\Bigg\{\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}\Bigg\}\Bigg\{r^{1/2}p^{-1}(1-2)^{-1/2}\Bigg\}$$
-From this posterior we obtain the full conditionals. First consider $p|r,X$:
+We will use a non-informative random variable, namely, Jeffreys' prior, giving us the following full conditionals: 
 $$p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$$
 We recognize this as the kernel of a beta distribution, namely
-
 $$p|r,X\sim\mathcal{B}eta(n\bar{X},nr+1/2).$$
@@
 $$p(r|p,X)\propto\bigg[\prod_{i=1}^n\Gamma(r+x_i)\bigg]\Gamma(r)^{-n}(1-p)^{nr}r^{1/2}$$
-This is not a recognized distribtion. So if we wish to make inferences on $r$ we must use a Metropolis algorithm to sample from it.
+This is not a recognized distribtion. So if we wish to make inferences on $r$ we must use a Metropolis algorithm to sample from it. For a full derivation see~\ref{a:dNB}.
 
@@
 This distribution does not factor nicely, so I will use the Metropolis algorithm to sample from it.
-Because this posterior does not suggest any obvious proposal distributions I will sample each independently from a normal distribution centered at $\theta^*$, and with a variance that is tuned to yeild an appropriate acceptance rate.
+Because this posterior does not suggest any obvious proposal distributions I will sample each iteratively from a truncated normal distribution centered at $\theta^*$, and with a variance that is tuned to yield an appropriate acceptance rate.
 
@@
 
+Because the posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, we were able to sample from this distribution directly.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $\lambda$.
+The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisnbdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $\lambda$, for both event types. The second column relays the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
+    \label{t:rPOIS}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$  \\ \hline
+    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
+    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=.45\textwidth]{figure/POIS_Density.pdf}
+\includegraphics[width=.45\textwidth]{figure/NB_Density.pdf}
+\caption{Posterior density plots for Poisson (left) and Negative Binomial (right) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:poisnbdensity}
+\end{figure}
+
 \subsection{Negative Binomial}
 \label{ss:rNBinom}
+
+As in the case of the Poisson distribution, the conditional posterior from our Negative Binomial likelihood and noninformative prior had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly.
+Meanwhile we took $r$ as known.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $p$.
+The results are summarized in Table~\ref{t:rNB}, while density plots can be found in Figure~\ref{f:poisnbdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $p$, for both event types. The second column relays the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
+    \label{t:rNB}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $p$  \\ \hline
+    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
+    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
+    \end{tabular}
+\end{table}
 
@@
 The results are summarized in Table~\ref{t:rZIP}.
-Convergance plots can be found in Figrue~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.
+Density plots can be found in Figure~\ref{f:zipfig}.
 
@@
 
-\begin{figure}[p]
+\begin{figure}
 \centering
-\includegraphics[width=.65\textwidth]{figure/ZIP_Flood_Conv.pdf}
-\caption{Cumulative mean plots for the posterior samples (ZIP model) of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.}
-\label{f:zipconv}
-\end{figure}
-
-\begin{figure}[p]
-\centering
+%\includegraphics[width=.65\textwidth]{figure/ZIP_Flood_Conv.pdf}\\
 \includegraphics[width=.65\textwidth]{figure/ZIP_Density.pdf}
-\caption{Posterior density plots (ZIP model) for flood parameters (in blue), and tornado parameters (black).
-The solid line represents the mean and the dashed lines indicate the 95\% confidence interval.}
-\label{f:zipdensity}
+\caption{
+%Top: Cumulative mean plots for the posterior samples (ZIP model) of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly. Bottom:
+Posterior density plots (ZIP model) for flood parameters (in blue), and tornado parameters (black). The solid line represents the mean and the dashed lines indicate the 95\% confidence interval.}
+\label{f:zipfig}
 \end{figure}
@@
 It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary.
-Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+%Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
 <<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
@@
 In the end $\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
-The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$.
+%The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$.
 The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.
@@
 @
-
 The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.
@@
 \end{table}
-\begin{figure}[p]
-	\centering
+%\begin{figure}[p]
+%	\centering
  <<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=1.75>>=
-B<-length(flood.zinb$sigma)
-Bt<-length(tornado.zinb$sigma)
-par(mfrow=c(1, 3))
-par(mgp=c(1.25,.5,0))
-par(mar=c(3.1,4.1,2.5,.1))
-plot(cumsum(flood.zinb$sigma)/1:B, 
-     main = expression(paste('Cum. Mean (Flood) ',sigma)),
-     ylab = 'Cumulative Mean',
-     type='l')
-plot(cumsum(flood.zinb$r)/1:B, 
-     main = expression(paste('Cum. Mean (Flood) ',r)),
-     ylab = 'Cumulative Mean',
-     type='l')
-plot(cumsum(flood.zinb$p)/1:B, 
-     type='l', 
-     main = expression(paste('Cum. Mean (Flood) ',p)),
-     ylab = 'Cumulative Mean')
+# B<-length(flood.zinb$sigma)
+# Bt<-length(tornado.zinb$sigma)
+# par(mfrow=c(1, 3))
+# par(mgp=c(1.25,.5,0))
+# par(mar=c(3.1,4.1,2.5,.1))
+# plot(cumsum(flood.zinb$sigma)/1:B, 
+#      main = expression(paste('Cum. Mean (Flood) ',sigma)),
+#      ylab = 'Cumulative Mean',
+#      type='l')
+# plot(cumsum(flood.zinb$r)/1:B, 
+#      main = expression(paste('Cum. Mean (Flood) ',r)),
+#      ylab = 'Cumulative Mean',
+#      type='l')
+# plot(cumsum(flood.zinb$p)/1:B, 
+#      type='l', 
+#      main = expression(paste('Cum. Mean (Flood) ',p)),
+#      ylab = 'Cumulative Mean')
 # acf(flood.zinb$sigma, 
@@
 @
-	\caption{Cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
-	\label{f:zinbconvergence}
-\end{figure}
+%	\caption{Cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
+%	\label{f:zinbconvergence}
+%\end{figure}
 
-\begin{figure}[p]
+\begin{figure}
 	\centering
@@
 @
-	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
+	\caption{Posterior density curves for each of the three variables (ZINB model). The flood posterior is shown in blue and the tornado posterior is shown in black.}
 	\label{f:zinbdensity}
@@
 \label{t:evalresults}
-\begin{tabular}{lcc}
+\begin{tabular}{lrr}
 \toprule
@@
 \midrule
-Poisson & P FLOOD DIC& P TORNADO DIC\\
-Negative Binomial & NB FLOOD DIC&NB TORNADO DIC\\
-ZIP & 10454.860&8727.647 \\
-ZINB & \Sexpr{flood.dic.zinb}&\Sexpr{tornado.dic.zinb}\\
+Poisson & 14237 & 17606\\
+Negative Binomial & 101548 & 26372\\
+ZIP & 10454 & 8727 \\
+ZINB & \Sexpr{as.integer(flood.dic.zinb)}&\Sexpr{as.integer(tornado.dic.zinb)}\\
 \bottomrule
@@
 This appendix will include details on the calculations required to derive our models.
+\subsection{Poisson Derivation}
+\label{a:dP}
+The likelihood for a sum of Poisson random variables can be written as follows:
+$$\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{X}}e^{-n\lambda}.$$ 
+Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
+$$\pi(\lambda)\propto\lambda^{1/2-1}e^{-0\cdot\lambda}.$$
+We recognize this as the kernel of an improper gamma distribution. Combining the likelihood and the prior distribtion yields the following posterior distribution:
+$$p(\lambda|X)=\lambda^{n\bar{X}+1/2-1}e^{-n\lambda}.$$
+We recognize this as the kernel of a gamma distribution, namely
+$$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$
+
+\subsection{Negatvie Binomial Derivation}
+\label{a:dNB}
+The likelihood for a sum of Negative Binomial random variables can be written as follows:
+$$\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}.$$
+Since we want the data to speak for itself, we will use a non-informative random variable, namely, Jeffreys' prior. For the Poisson distribtion this is given as follows:
+$$\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}.$$
+Combining the likelihood and the prior distribtion yields the following posterior distribution:
+$$p(r,p|X)=\Bigg\{\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}\Bigg\}\Bigg\{r^{1/2}p^{-1}(1-2)^{-1/2}\Bigg\}$$
+From this posterior we obtain the full conditionals. First consider $p|r,X$:
+$$p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$$
+We recognize this as the kernel of a beta distribution, namely
+
+$$p|r,X\sim\mathcal{B}eta(n\bar{X},nr+1/2).$$
+Next consider $r|p,X$:
+$$p(r|p,X)\propto\bigg[\prod_{i=1}^n\Gamma(r+x_i)\bigg]\Gamma(r)^{-n}(1-p)^{nr}r^{1/2}$$
+
 \subsection{Zero Inflated Poisson Derivation}

### 2017-05-11 paper_edits.tex
Edit for length

--- 2017-05-10 paper.tex
@@
 This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
-The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
-For this project we focused on the deaths directly attributable to the event.
+We chose to use a subset of that data covering the years 1996 through 2016.
+The data coms from the NOAA website \cite{dataset}.
+For this project we focused on the deaths directly attributable to a type of weather event.
+In the interest of time, we limited our models to two types of severe weather, Tornados and Flash Floods, reponsible for the second and third most direct deaths among weather types.
 Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
@@
 
-\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
+\noindent Bayarri, Berger, and Datta (2008) suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior
 
@@
 
+Because the posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, we were able to sample from this distribution directly.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $\lambda$.
+The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $\lambda$, for both event types. The second column relays the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
+    \label{t:rPOIS}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$  \\ \hline
+    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
+    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/POIS_Density.pdf}
+\caption{Posterior density plots (Poisson model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:poisdensity}
+\end{figure}
+
 \subsection{Negative Binomial}
@@
 
+As in the case of the Poisson distribution, the conditional posterior from our Negative Binomial likelihood and noninformative prior had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly.
+Meanwhile we took $r$ as known.
+For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $p$.
+The results are summarized in Table~\ref{t:rNB}, while density plots can be found in Figure~\ref{f:nbdensity}.
+
+\begin{table}
+    \centering
+    \caption{Posterior distributions of $p$, for both event types. The second column relays the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
+    \label{t:rNB}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $p$  \\ \hline
+    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
+    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/NB_Density.pdf}
+\caption{Posterior density plots (Negative Binomial model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:nbdensity}
+\end{figure}
+
 \subsection{Zero Inflated Poisson}
 \label{ss:rZiPoisson}
+
+In tuning the parameters of the two models, slightly different values were chosen for the two different event types.
+The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods.
+The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model.
+
+For each variable of interest, 20,000 samples were taken.
+As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in.
+The results are summarized in Table~\ref{t:rZIP}.
+Convergance plots can be found in Figure~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.
+
+\begin{table}
+    \centering
+    \caption{The posterior distributions of $\lambda$ and p, for both event types.
+The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
+The third does the same for p.}
+    \label{t:rZIP}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$ & p  \\ \hline
+    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
+    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/ZIP_Flood_Conv.pdf}
+\caption{Cumulative mean plots for the posterior samples (ZIP model) of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.}
+\label{f:zipconv}
+\end{figure}
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/ZIP_Density.pdf}
+\caption{Posterior density plots (ZIP model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the mean and the dashed lines indicate the 95\% confidence interval.}
+\label{f:zipdensity}
+\end{figure}
 
@@
 This model suffered severely from the curse of dimensionality.
-It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary.
-Autocorrelation and running mean plots for both the Tornado and the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary.
+Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
 
-In the end $650000$ samples were drawn originally for each model, with a burn in of $20000$ and one of every $70$ samples was retained, resulting in a final sample size of $9000$.
-The flood model had an acceptance rate of $0.193$ and the tornado model had an acceptance rate of $0.193$.
-The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
+In the end $1030000$ samples were drawn for each model, with a burn in of $30000$ and one of every $100$ samples was retained, resulting in a final sample size of $10000$.
+The flood model had an acceptance rate of $0.195$ and the tornado model had an acceptance rate of $0.195$.
+The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.
 
-The DIC for floods and tornadoes respectively were $10804$ and $7166$. I was unable to calculate WIACs because it was too computationally expensive.
+The DIC for floods and tornadoes respectively were $10804$ and $7166$.
 \begin{table}
@@
 \midrule
-$\sigma$ & 0.305 &
-  0.285&
-  (0.014, 
-    0.712)&
-  0.29 &
-  0.267&
+$\sigma$ & 0.307 &
+  0.282&
+  (0.013, 
+    0.716)&
+  0.299 &
+  0.276&
   (0.012, 
-    0.682)\\
-$p$ &  0.555 &
+    0.702)\\
+$p$ &  0.554 &
   0.555&
-  (0.519, 
-    0.591)&
+  (0.518, 
+    0.589)&
   0.862 &
@@
     0.882)\\
-$r$ &  0.023 &
+$r$ &  0.024 &
   0.02&
@@
   (0.009, 
-    0.03)\\
+    0.033)\\
 \bottomrule
@@
 \end{table}
-\begin{figure}
+\begin{figure}[p]
 	\centering
@@
 \end{knitrout}
-	\caption{Autocorrelation plots, index plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood and the tornado data.}
+	\caption{Cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
 	\label{f:zinbconvergence}
@@
 
-\begin{figure}
+\begin{figure}[p]
 	\centering
@@
 \label{s:discussion}
+The goal of this analysis was to determine which of the four models was the best fit for the data.
+In order to make this determination the DIC was calculated for each model on both tornados and flash floods.
+The results are summarized in Table~\ref{t:evalresults}.
 
+\begin{table}
+\centering
+\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
+\label{t:evalresults}
+\begin{tabular}{lcc}
+\toprule
+Model & DIC Flash Flood & DIC Tornado\\
+\midrule
+Poisson & 14237 & 17606\\
+Negative Binomial & 101548 & 26372\\
+ZIP & 10454 & 8727 \\
+ZINB & 10804.104&7166.309\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+In general, models with smaller DIC are preferred to models with larger DIC.
+From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados.
+Thus for estimating the rate of leathality of Flash Floods, Tornados, and possibly other severe weather, we should use a Zero Inflated Poisson model to fit the data.
+
+\clearpage
 \begin{thebibliography}{2}
@@
 \end{knitrout}
+\subsection{Zero Inflated Poisson code}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(dplyr)}
+\hlstd{tornado} \hlkwb{<-} \hlstd{data} \hlopt{%>%} \hlkwd{filter}\hlstd{(EVENT_TYPE} \hlopt{==} \hlstr{"Tornado"}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+\hlstd{flood} \hlkwb{<-} \hlstd{data} \hlopt{%>%} \hlkwd{filter}\hlstd{(EVENT_TYPE} \hlopt{==} \hlstr{"Flash Flood"}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+
+\hlcom{# Precompute values to use in posterior function}
+\hlstd{zeros} \hlkwb{<-} \hlstd{tornado} \hlopt{%>%} \hlkwd{filter}\hlstd{(DEATHS_DIRECT} \hlopt{==} \hlnum{0}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+\hlstd{ones} \hlkwb{<-} \hlstd{tornado} \hlopt{%>%} \hlkwd{filter}\hlstd{(DEATHS_DIRECT} \hlopt{!=} \hlnum{0}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+\hlstd{n} \hlkwb{<-} \hlkwd{nrow}\hlstd{(ones)}
+\hlstd{t0_exp} \hlkwb{<-} \hlstd{(zeros)}\hlopt{/}\hlkwd{factorial}\hlstd{(zeros)}
+\hlstd{l_t1_exp} \hlkwb{<-} \hlstd{ones}\hlopt{-}\hlnum{1}\hlopt{/}\hlnum{2}
+
+\hlcom{# Define Posterior Functions}
+\hlstd{log_lam} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{y}\hlstd{,} \hlkwc{p}\hlstd{,} \hlkwc{lambda}\hlstd{)\{}
+  \hlstd{t0} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{log}\hlstd{(p} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}\hlopt{*}\hlkwd{exp}\hlstd{(}\hlopt{-}\hlnum{1}\hlopt{*}\hlstd{lambda)}\hlopt{*}\hlstd{(lambda}\hlopt{+}\hlnum{2e-10}\hlstd{)}\hlopt{^}\hlstd{t0_exp))}
+  \hlstd{t1} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{log}\hlstd{(}\hlkwd{exp}\hlstd{(}\hlopt{-}\hlnum{1}\hlopt{*}\hlstd{lambda)}\hlopt{*}\hlstd{lambda}\hlopt{^}\hlstd{l_t1_exp))}
+  \hlkwd{return}\hlstd{(t0}\hlopt{+}\hlstd{t1)}
+\hlstd{\}}
+
+\hlstd{log_p} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{y}\hlstd{,} \hlkwc{p}\hlstd{,} \hlkwc{lambda}\hlstd{)\{}
+  \hlstd{t0} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{log}\hlstd{(p} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p)}\hlopt{*}\hlkwd{exp}\hlstd{(}\hlopt{-}\hlnum{1}\hlopt{*}\hlstd{lambda)}\hlopt{*}\hlstd{(lambda}\hlopt{+}\hlnum{2e-10}\hlstd{)}\hlopt{^}\hlstd{t0_exp))}
+  \hlstd{t1_vec} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{1}\hlopt{-}\hlstd{p, n)}
+  \hlstd{t1} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{log}\hlstd{(t1_vec))}
+  \hlkwd{return}\hlstd{(t0}\hlopt{+}\hlstd{t1)}
+\hlstd{\}}
+
+\hlcom{### Sampler for Tornado}
+\hlkwd{set.seed}\hlstd{(}\hlnum{2011}\hlstd{)}
+
+\hlstd{l_last} \hlkwb{<-} \hlnum{1}
+\hlstd{p_last} \hlkwb{<-} \hlnum{.98}
+
+\hlstd{B} \hlkwb{<-} \hlnum{20000}
+\hlstd{lambdas} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+\hlstd{ps} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+
+\hlstd{l_accept} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{FALSE}\hlstd{, B)}
+\hlstd{p_accept} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{FALSE}\hlstd{, B)}
+
+\hlkwa{for}\hlstd{(t} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{B)\{}
+  \hlstd{lambda} \hlkwb{<-} \hlkwd{rgamma}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{3}\hlstd{,} \hlnum{2}\hlstd{)}
+  \hlstd{p} \hlkwb{<-} \hlkwd{rbeta}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{1940}\hlstd{,} \hlnum{60}\hlstd{)}
+  \hlstd{ro_lam} \hlkwb{<-} \hlkwd{log_lam}\hlstd{(tornado, p_last, lambda)} \hlopt{-} \hlkwd{log}\hlstd{(}\hlkwd{dgamma}\hlstd{(lambda,} \hlnum{3}\hlstd{,} \hlnum{2}\hlstd{))} \hlopt{-}
+    \hlkwd{log_lam}\hlstd{(tornado, p_last, l_last)} \hlopt{+} \hlkwd{log}\hlstd{(}\hlkwd{dgamma}\hlstd{(l_last,} \hlnum{3}\hlstd{,} \hlnum{2}\hlstd{))}
+  \hlstd{ro_p} \hlkwb{<-} \hlkwd{log_p}\hlstd{(tornado, p, l_last)} \hlopt{-} \hlkwd{log}\hlstd{(}\hlkwd{dbeta}\hlstd{(p,} \hlnum{1940}\hlstd{,} \hlnum{60}\hlstd{))} \hlopt{-}
+    \hlkwd{log_p}\hlstd{(tornado, p_last, l_last)} \hlopt{+} \hlkwd{log}\hlstd{(}\hlkwd{dbeta}\hlstd{(p_last,} \hlnum{1940}\hlstd{,} \hlnum{60}\hlstd{))}
+  \hlstd{U_lambda} \hlkwb{<-} \hlkwd{runif}\hlstd{(}\hlnum{1}\hlstd{)}
+  \hlstd{U_p} \hlkwb{<-} \hlkwd{runif}\hlstd{(}\hlnum{1}\hlstd{)}
+  \hlkwa{if}\hlstd{(}\hlkwd{log}\hlstd{(U_lambda)} \hlopt{<} \hlkwd{min}\hlstd{(}\hlnum{0}\hlstd{, ro_lam))\{}
+    \hlkwd{print}\hlstd{(}\hlnum{TRUE}\hlstd{)}
+    \hlstd{lambdas[t]} \hlkwb{<-} \hlstd{lambda}
+    \hlstd{lambda_last} \hlkwb{<-} \hlstd{lambda}
+    \hlstd{l_accept[t]} \hlkwb{<-} \hlnum{TRUE}
+  \hlstd{\}} \hlkwa{else}\hlstd{\{}
+    \hlkwd{print}\hlstd{(}\hlnum{FALSE}\hlstd{)}
+    \hlstd{lambdas[t]} \hlkwb{<-} \hlstd{lambda_last}
+  \hlstd{\}}
+
+  \hlkwa{if}\hlstd{(}\hlkwd{log}\hlstd{(U_p)} \hlopt{<} \hlkwd{min}\hlstd{(}\hlnum{0}\hlstd{, ro_p))\{}
+    \hlkwd{print}\hlstd{(}\hlnum{TRUE}\hlstd{)}
+    \hlstd{ps[t]} \hlkwb{<-} \hlstd{p}
+    \hlstd{p_last} \hlkwb{<-} \hlstd{p}
+    \hlstd{p_accept[t]} \hlkwb{<-} \hlnum{TRUE}
+  \hlstd{\}} \hlkwa{else}\hlstd{\{}
+    \hlkwd{print}\hlstd{(}\hlnum{FALSE}\hlstd{)}
+    \hlstd{ps[t]} \hlkwb{<-} \hlstd{p_last}
+  \hlstd{\}}
+\hlstd{\}}
+
+\hlcom{# Do the same for flash floods}
+
+\hlcom{# Update values for likelihood function}
+\hlstd{zeros} \hlkwb{<-} \hlstd{flood} \hlopt{%>%} \hlkwd{filter}\hlstd{(DEATHS_DIRECT} \hlopt{==} \hlnum{0}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+\hlstd{ones} \hlkwb{<-} \hlstd{flood} \hlopt{%>%} \hlkwd{filter}\hlstd{(DEATHS_DIRECT} \hlopt{!=} \hlnum{0}\hlstd{)} \hlopt{%>%} \hlkwd{select}\hlstd{(DEATHS_DIRECT)}
+\hlstd{n} \hlkwb{<-} \hlkwd{nrow}\hlstd{(ones)}
+\hlstd{t0_exp} \hlkwb{<-} \hlstd{(zeros)}\hlopt{/}\hlkwd{factorial}\hlstd{(zeros)}
+\hlstd{l_t1_exp} \hlkwb{<-} \hlstd{ones}\hlopt{-}\hlnum{1}\hlopt{/}\hlnum{2}
+
+\hlcom{# Sampler:}
+\hlkwd{set.seed}\hlstd{(}\hlnum{2012}\hlstd{)}
+\hlstd{l_last} \hlkwb{<-} \hlnum{1}
+\hlstd{p_last} \hlkwb{<-} \hlnum{.98}
+\hlstd{B} \hlkwb{<-} \hlnum{20000}
+\hlstd{lambdas} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+\hlstd{ps} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{0}\hlstd{, B)}
+\hlstd{l_accept} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{FALSE}\hlstd{, B)}
+\hlstd{p_accept} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{FALSE}\hlstd{, B)}
+\hlkwa{for}\hlstd{(t} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{B)\{}
+  \hlkwd{print}\hlstd{(}\hlkwd{paste0}\hlstd{(}\hlstr{"t = "}\hlstd{,} \hlkwd{as.character}\hlstd{(t)))}
+  \hlstd{lambda} \hlkwb{<-} \hlkwd{rgamma}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{2}\hlstd{)}
+  \hlkwd{print}\hlstd{(}\hlkwd{paste0}\hlstd{(}\hlstr{"lambda = "}\hlstd{,} \hlkwd{as.character}\hlstd{(lambda)))}
+  \hlstd{p} \hlkwb{<-} \hlkwd{rbeta}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2945}\hlstd{,} \hlnum{55}\hlstd{)}
+  \hlkwd{print}\hlstd{(}\hlkwd{paste0}\hlstd{(}\hlstr{"p = "}\hlstd{,} \hlkwd{as.character}\hlstd{(p)))}
+  \hlstd{ro_lam} \hlkwb{<-} \hlkwd{log_lam}\hlstd{(flood, p_last, lambda)} \hlopt{-} \hlkwd{log}\hlstd{(}\hlkwd{dgamma}\hlstd{(lambda,} \hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))} \hlopt{-}
+    \hlkwd{log_lam}\hlstd{(flood, p_last, l_last)} \hlopt{+} \hlkwd{log}\hlstd{(}\hlkwd{dgamma}\hlstd{(l_last,} \hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+  \hlstd{ro_p} \hlkwb{<-} \hlkwd{log_p}\hlstd{(flood, p, l_last)} \hlopt{-} \hlkwd{log}\hlstd{(}\hlkwd{dbeta}\hlstd{(p,} \hlnum{2945}\hlstd{,} \hlnum{55}\hlstd{))} \hlopt{-}
+    \hlkwd{log_p}\hlstd{(flood, p_last, l_last)} \hlopt{+} \hlkwd{log}\hlstd{(}\hlkwd{dbeta}\hlstd{(p_last,} \hlnum{2945}\hlstd{,} \hlnum{55}\hlstd{))}
+  \hlstd{U_lambda} \hlkwb{<-} \hlkwd{runif}\hlstd{(}\hlnum{1}\hlstd{)}
+  \hlstd{U_p} \hlkwb{<-} \hlkwd{runif}\hlstd{(}\hlnum{1}\hlstd{)}
+  \hlkwa{if}\hlstd{(}\hlkwd{log}\hlstd{(U_lambda)} \hlopt{<} \hlkwd{min}\hlstd{(}\hlnum{0}\hlstd{, ro_lam))\{}
+    \hlkwd{print}\hlstd{(}\hlnum{TRUE}\hlstd{)}
+    \hlstd{lambdas[t]} \hlkwb{<-} \hlstd{lambda}
+    \hlstd{lambda_last} \hlkwb{<-} \hlstd{lambda}
+    \hlstd{l_accept[t]} \hlkwb{<-} \hlnum{TRUE}
+  \hlstd{\}} \hlkwa{else}\hlstd{\{}
+    \hlkwd{print}\hlstd{(}\hlnum{FALSE}\hlstd{)}
+    \hlstd{lambdas[t]} \hlkwb{<-} \hlstd{lambda_last}
+  \hlstd{\}}
+
+  \hlkwa{if}\hlstd{(}\hlkwd{log}\hlstd{(U_p)} \hlopt{<} \hlkwd{min}\hlstd{(}\hlnum{0}\hlstd{, ro_p))\{}
+    \hlkwd{print}\hlstd{(}\hlnum{TRUE}\hlstd{)}
+    \hlstd{ps[t]} \hlkwb{<-} \hlstd{p}
+    \hlstd{p_last} \hlkwb{<-} \hlstd{p}
+    \hlstd{p_accept[t]} \hlkwb{<-} \hlnum{TRUE}
+  \hlstd{\}} \hlkwa{else}\hlstd{\{}
+    \hlkwd{print}\hlstd{(}\hlnum{FALSE}\hlstd{)}
+    \hlstd{ps[t]} \hlkwb{<-} \hlstd{p_last}
+  \hlstd{\}}
+\hlstd{\}}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+\subsection{Poisson code}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{set.seed}\hlstd{(}\hlnum{05112017}\hlstd{)}
+
+\hlcom{# --- model evaluation function -----------------------------------------------------------}
+\hlstd{DIC} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{y}\hlstd{,} \hlkwc{theta}\hlstd{,} \hlkwc{loglik}\hlstd{) \{}
+  \hlstd{theta_hat} \hlkwb{=} \hlkwd{mean}\hlstd{(theta)} \hlcom{#theta_hat = apply(theta, 2, mean)}
+  \hlstd{L} \hlkwb{=} \hlkwd{loglik}\hlstd{(y, theta_hat)}
+  \hlstd{S} \hlkwb{=} \hlkwd{length}\hlstd{(theta)} \hlcom{#S = nrow(theta) #S = number of iterations}
+  \hlstd{llSum} \hlkwb{=} \hlnum{0}
+  \hlkwa{for} \hlstd{(s} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{S) \{}
+    \hlstd{theta_s} \hlkwb{=} \hlstd{theta[s]}
+    \hlstd{llSum} \hlkwb{=} \hlstd{llSum} \hlopt{+} \hlkwd{loglik}\hlstd{(y, theta_s)}
+  \hlstd{\}}
+  \hlstd{P} \hlkwb{=} \hlnum{2} \hlopt{*} \hlstd{(L} \hlopt{-} \hlstd{(}\hlnum{1} \hlopt{/} \hlstd{S} \hlopt{*} \hlstd{llSum))}
+  \hlstd{DIC} \hlkwb{=} \hlopt{-} \hlnum{2} \hlopt{*} \hlstd{(L} \hlopt{-} \hlstd{P)}
+  \hlkwd{return}\hlstd{(DIC)}
+\hlstd{\}}
+
+\hlcom{# --- poisson log-likelihood ---------------------------------------------------------}
+\hlstd{pois_ll} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{,} \hlkwc{lambda}\hlstd{) \{}
+  \hlstd{ll} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{dpois}\hlstd{(x, lambda,} \hlkwc{log} \hlstd{=} \hlnum{TRUE}\hlstd{))}
+  \hlkwd{return}\hlstd{(ll)}
+\hlstd{\}}
+
+\hlcom{# --- poisson model -----------------------------------------------------------------}
+\hlstd{pois_fit} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{data}\hlstd{,} \hlkwc{nsim}\hlstd{) \{}
+  \hlstd{x} \hlkwb{<-} \hlstd{data}
+  \hlstd{xbar} \hlkwb{<-} \hlkwd{mean}\hlstd{(x)}
+  \hlstd{n} \hlkwb{<-} \hlkwd{length}\hlstd{(x)}
+  \hlstd{lambda} \hlkwb{<-} \hlkwd{rgamma}\hlstd{(nsim, n} \hlopt{*} \hlstd{xbar} \hlopt{+} \hlnum{1} \hlopt{/}\hlnum{2}\hlstd{, n)}
+  \hlkwd{return}\hlstd{(lambda)}
+\hlstd{\}}
+
+\hlcom{# --- results - ---------------------------------------------------------------------------}
+\hlstd{lambda_t} \hlkwb{<-} \hlkwd{pois_fit}\hlstd{(}\hlkwc{data} \hlstd{= TORNADO}\hlopt{$}\hlstd{DEATHS,} \hlkwc{nsim} \hlstd{=} \hlnum{10000}\hlstd{)}
+\hlstd{lambda_ff} \hlkwb{<-} \hlkwd{pois_fit}\hlstd{(}\hlkwc{data} \hlstd{= FLASH_FLOOD}\hlopt{$}\hlstd{DEATHS,} \hlkwc{nsim} \hlstd{=} \hlnum{10000}\hlstd{)}
+
+\hlkwd{quantile}\hlstd{(lambda_t,} \hlkwd{c}\hlstd{(}\hlnum{0.025}\hlstd{,} \hlnum{0.5}\hlstd{,} \hlnum{0.975}\hlstd{))}
+\hlkwd{quantile}\hlstd{(lambda_ff,} \hlkwd{c}\hlstd{(}\hlnum{0.025}\hlstd{,} \hlnum{0.5}\hlstd{,} \hlnum{0.975}\hlstd{))}
+
+\hlkwd{DIC}\hlstd{(TORNADO}\hlopt{$}\hlstd{DEATHS, lambda_t, pois_ll)}
+\hlkwd{DIC}\hlstd{(FLASH_FLOOD}\hlopt{$}\hlstd{DEATHS, lambda_ff, pois_ll)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\subsection{Negative Binomial code}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{# --- negative binomial log-likelihood ---------------------------------------------------------}
+\hlstd{nb_ll} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{,} \hlkwc{p}\hlstd{) \{}
+  \hlstd{ll} \hlkwb{<-} \hlkwd{sum}\hlstd{(}\hlkwd{dnbinom}\hlstd{(}\hlkwc{x} \hlstd{= x,} \hlkwc{prob} \hlstd{= p,} \hlkwc{size} \hlstd{=} \hlnum{1} \hlopt{/} \hlnum{4}\hlstd{,} \hlkwc{log} \hlstd{=} \hlnum{TRUE}\hlstd{))}
+  \hlkwd{return}\hlstd{(ll)}
+\hlstd{\}}
+
+\hlcom{# --- negative binomial model -----------------------------------------------------------------}
+\hlstd{nb_fit} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{data}\hlstd{,} \hlkwc{nsim}\hlstd{) \{}
+  \hlstd{x} \hlkwb{<-} \hlstd{data}
+  \hlstd{xbar} \hlkwb{<-} \hlkwd{mean}\hlstd{(x)}
+  \hlstd{n} \hlkwb{<-} \hlkwd{length}\hlstd{(x)}
+  \hlstd{r} \hlkwb{<-} \hlnum{1} \hlopt{/} \hlnum{4}
+  \hlstd{p} \hlkwb{<-} \hlkwd{rbeta}\hlstd{(nsim, n} \hlopt{*} \hlstd{xbar, n} \hlopt{*} \hlstd{r} \hlopt{+} \hlnum{1} \hlopt{/} \hlnum{2}\hlstd{)}
+  \hlkwd{return}\hlstd{(p)}
+\hlstd{\}}
+
+\hlcom{# --- results - ---------------------------------------------------------------------------}
+\hlstd{p_t} \hlkwb{<-} \hlkwd{nb_fit}\hlstd{(}\hlkwc{data} \hlstd{= TORNADO}\hlopt{$}\hlstd{DEATHS,} \hlkwc{nsim} \hlstd{=} \hlnum{10000}\hlstd{)}
+\hlstd{p_ff} \hlkwb{<-} \hlkwd{nb_fit}\hlstd{(}\hlkwc{data} \hlstd{= FLASH_FLOOD}\hlopt{$}\hlstd{DEATHS,} \hlkwc{nsim} \hlstd{=} \hlnum{10000}\hlstd{)}
+
+\hlcom{# --- credible intervals ----------------------------------------------------------------------}
+\hlkwd{quantile}\hlstd{(p_t,} \hlkwd{c}\hlstd{(}\hlnum{0.025}\hlstd{,} \hlnum{0.5}\hlstd{,} \hlnum{0.975}\hlstd{))}
+\hlkwd{quantile}\hlstd{(p_ff,} \hlkwd{c}\hlstd{(}\hlnum{0.025}\hlstd{,} \hlnum{0.5}\hlstd{,} \hlnum{0.975}\hlstd{))}
+
+\hlcom{# --- dic ----------------------------------------------------------------------}
+\hlkwd{DIC}\hlstd{(TORNADO}\hlopt{$}\hlstd{DEATHS, p_t, nb_ll)}
+\hlkwd{DIC}\hlstd{(FLASH_FLOOD}\hlopt{$}\hlstd{DEATHS, p_ff, nb_ll)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 

### 2017-05-11 presentation.tex
Re-Ordered ZIP Slides
Added POIS, NB models and DIC table

@@
 
+	\section{Poisson Model}
+	\begin{frame}
+		\frametitle{Poisson Model}
+		% Slide one
+		The Poisson model has one parameter, $\lambda$ that represents the expected number of occurrences of the event of interest. For a single random variable $x$, the probability density is:
+$$p(x|\lambda)=\frac{\lambda^xe^{-\lambda}}{x!}.$$
+	\end{frame}
+	\begin{frame}
+		\frametitle{Poisson Model}
+		% Slide two
+		The Poisson likelihood can be written as follows:
+$$\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{X}}e^{-n\lambda}.$$  Since we want the data to speak for itself we use Jeffreys' prior, $$\pi(\lambda)\propto\lambda^{1/2-1}e^{-0\cdot\lambda}.$$ Thus the posterior is: $$p(\lambda|X)=\lambda^{n\bar{X}+1/2-1}e^{-n\lambda}.$$
+We recognize this as the kernel of a gamma distribution, namely
+$$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$
+
+	\end{frame}
+	\begin{frame}
+		\frametitle{Poisson Model}
+		% Slide three
+		\begin{table}
+    \centering
+    \caption{Posterior distributions of $\lambda$, for both event types.
+From 10,000 samples for each variable we obtained the following values: the second column is the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
+    \label{t:rPOIS}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $\lambda$  \\ \hline
+    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
+    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
+    \end{tabular}
+\end{table}
+	\end{frame}	
+		\begin{frame}
+		\frametitle{Poisson Model}
+		% Slide four
+		\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/POIS_Density.pdf}
+\caption{Posterior density plots (Poisson model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:poisdensity}
+\end{figure}
+	\end{frame}
+
 	\section{Negative Binomial Model}
@@
 		% Slide one
+		The Negative Binomial model has two parameters, $r,p$, we are mostly interested in $p$. For a single random variable $x$, the probability density is:
+$$p(X|r,p)=\frac{\Gamma(r+x)}{\Gamma(r)x!}p^{x}(1-p)^{r}.$$
 	\end{frame}
@@
 		% Slide two
-	\end{frame}
-
-	\section{Poisson Model}
-	\begin{frame}
-		\frametitle{Poisson Model}
-		% Slide one
+		The likelihood of the Negative Binomial can be written as:
+$$\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{X}}(1-p)^{nr}.$$
+Since we want the data to speak for itself, we will use Jeffreys' prior. For the Negative Binomial distribution this is:
+$$\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}.$$
+We are interested in the conditional posterior of $p|r,X$ given by:
+$$p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$$ 
+This is the kernel of a beta distribution, namely
+$$p|r,X\sim\mathcal{B}eta(n\bar{X},nr+1/2).$$
 	\end{frame}
 		\begin{frame}
-		\frametitle{Poisson Model}
-		% Slide two
+		\frametitle{Negative Binomial}
+		% Slide three
+		\begin{table}
+    \centering
+    \caption{Posterior distributions of $p$, for both event types.
+From 10,000 samples for each variable we obtained the following values: the second column is the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
+    \label{t:rNB}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Event Type & $p$  \\ \hline
+    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
+    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
+    \end{tabular}
+\end{table}
 	\end{frame}
-
+		\begin{frame}
+		\frametitle{Negative Binomial}
+		% Slide four
+		\begin{figure}[p]
+\centering
+\includegraphics[width=.65\textwidth]{figure/NB_Density.pdf}
+\caption{Posterior density plots (Negative Binomial model) for flood parameters (in blue), and tornado parameters (black).
+The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
+\label{f:nbdensity}
+\end{figure}
+	\end{frame}
 	\section{Zero Inflated Poisson Model}
@@
 		\frametitle{Zero Inflated Poisson Model}
+		Convergence of parameters (flood model):
+		\centering
+		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
+	\end{frame}
+	\begin{frame}
+		\frametitle{Zero Inflated Poisson Model}
 		\begin{table}
@@
 \end{table}
-	\end{frame}
-	\begin{frame}
-		\frametitle{Zero Inflated Poisson Model}
-		Convergence of parameters (flood model):
-		\centering
-		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
 	\end{frame}
@@
 		% Waic dic table goes here
+		\begin{table}
+\centering
+\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
+\label{t:evalresults}
+\begin{tabular}{lcc}
+\toprule
+Model & DIC Flash Flood & DIC Tornado\\
+\midrule
+Poisson & 14237 & 17606\\
+Negative Binomial & 101548 & 26372\\
+ZIP & 10454 & 8727 \\
+ZINB & 10804.104&7166.309\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Models with smaller DIC are preferred.
+From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados.
+Thus for estimating the rate of lethality of Flash Floods, Tornados, we should use a Zero Inflated Poisson model.
 	\end{frame}
 	\begin{frame}
-		\frametitle{Conclusions: Interpreting the Best Model}
-		% Waic dic table goes here
-	\end{frame}
-	\begin{frame}
+
 		\frametitle{Conclusions: Future Work}
@@
 	\end{frame}
-
 \end{document}

### 2017-05-11 paper.Rnw
under length! Yay!

@@
 \label{s:intro}
-This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
-We chose to use a subset of that data covering the years 1996 through 2016.
-The data coms from the NOAA website \cite{dataset}.
-For this project we focused on the deaths directly attributable to a type of weather event.
-In the interest of time, we considered two types of severe weather, Tornados and Flash Floods, reponsible for the second and third most direct deaths among weather types.
+Tornados and Flash Floods are reponsible for the second and third most direct deaths among weather types in the United States.
 Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
-In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.
+NOAA's website has a database of weather events and the death toll available on their website \cite{dataset}.
+For this study we looked at events from 1996-2016.
+
+The death toll from a severe weather event is a form of `count' data. 
+The most common model used for count data is the Poisson distribution. 
+However for some natural events, the negative binomial model is a better fit, because it is more flexible and allows for a difference between mean and variance.
+
+Fortunately, the vast majority of severe weather events in the United States involve no deaths, so to account for the possibility of structural zeros, therefore we also considered zero inflated variants.
+These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$.
+We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.
+
 The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models.
@@
 \label{s:methods}
-The deaths attributed to a severe weather event is `count' data. 
-The most common model used for count data is the Poisson distribution. 
-However for some weather events, the negative binomial model is a better fit, because the Poisson distribution assumes that the events being counted occur independently.
-
-Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants.
-These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$.
-We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.
-
 \subsection{Poisson}
@@
 $$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)$$
-This distribution does not factor nicely, so I will use the Metropolis algorithm to sample from it.
-Because this posterior does not suggest any obvious proposal distributions I will sample each iteratively from a truncated normal distribution centered at $\theta^*$, and with a variance that is tuned to yield an appropriate acceptance rate.
+This distribution does not factor, so I will use the Metropolis-Hastings algorithm to sample from it.
+I will use a truncated normal distribution centered at $\theta^*$ proposal density, and tune the variance for efficient convergence.
 
@@
 
-Because the posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, we were able to sample from this distribution directly.
-For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $\lambda$.
+The posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, therefore we were able to sample from this distribution directly.
+For both event types, 10,000 samples were taken, to simulate the variable of interest, $\lambda$.
 The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisnbdensity}.
@@
     \hline
-    Event Type & $\lambda$  \\ \hline
-    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
-    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
+    Event Type &  Tornado &  Flash Flood \\ \hline
+    $\lambda$ & 0.0592 (0.0564, 0.0620)  & 0.0179 (0.0170, 0.0189)\\
+    \hline
     \end{tabular}
@@
 \centering
-\includegraphics[width=.45\textwidth]{figure/POIS_Density.pdf}
-\includegraphics[width=.45\textwidth]{figure/NB_Density.pdf}
+\includegraphics[width=.4\textwidth]{figure/POIS_Density.pdf}
+\includegraphics[width=.4\textwidth]{figure/NB_Density.pdf}
 \caption{Posterior density plots for Poisson (left) and Negative Binomial (right) for flood parameters (in blue), and tornado parameters (black).
@@
 
-As in the case of the Poisson distribution, the conditional posterior from our Negative Binomial likelihood and noninformative prior had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly.
-Meanwhile we took $r$ as known.
-For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $p$.
-The results are summarized in Table~\ref{t:rNB}, while density plots can be found in Figure~\ref{f:poisnbdensity}.
-
-\begin{table}
-    \centering
-    \caption{Posterior distributions of $p$, for both event types. The second column relays the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
-    \label{t:rNB}
-    \begin{tabular}{| l | l | l |}
-    \hline
-    Event Type & $p$  \\ \hline
-    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
-    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
-    \end{tabular}
-\end{table}
-
+The conditional posterior from our Negative Binomial likelihood and noninformative prior also had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly.
+We took $r$ as known.
+For both event types, 10,000 samples were taken.
+The results are summarized in Table~\ref{t:rNB}, and density plots can be found in Figure~\ref{f:poisnbdensity}.
 \subsection{Zero Inflated Poisson}
 \label{ss:rZiPoisson}
-
-In tuning the parameters of the two models, slightly different values were chosen for the two different event types.
 The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods.
 The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model.
-
-For each variable of interest, 20,000 samples were taken.
+For each variable, 20,000 samples were taken.
 As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in.
@@
     \centering
-    \caption{The posterior distributions of $\lambda$ and p, for both event types.
-The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses.
-The third does the same for p.}
+      \caption{The posterior distributions of $\lambda$ and p, for both event types. The second column relays the mean and 95 percent credible interval in parentheses. The third does the same for p.}
       \label{t:rZIP}
@@
 <<zinbparams, echo=FALSE, cache=TRUE>>=
-samplesize <- 1030000
+samplesize <- 2030000
 burnin <- 30000
-thin <- 100
+thin <- 250
 r.var=.00002
@@
 @
-The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution.
-For the source code of the sampler, please see \ref{a:cZINB}.
-
-This model suffered severely from the curse of dimensionality.
-It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary.
-%Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
 <<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
@@
 @
-In the end $\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
+The ZINB model was fit using the Metropolis-Hastings algorithm with a truncated multivariate normal proposal distribution.
+For the source code of the sampler, please see \ref{a:cZINB}.
+Because it was fitting 3 parameters simultaniously this model was slow to converge, and had high autocorrelation, requiring agressive thinning.
+%Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}.
+$\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, for a final sample size of $\Sexpr{length(flood.zinb$sigma)}$.
 %The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$.
-The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.
+The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}.
+The densities for $p$ and $r$ are readily interpretable:  $p$ can be understood to measure dangerousness and $r$ can be thought of as duration.
+This implies floods are longer lasting than tornadoes, but less dangerous.
 <<evaluation, echo=FALSE, cache=FALSE>>=
@@
 @
-The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.
 \begin{table}
@@
 \label{t:ZINBparams}
+\scalebox{.8}{
 \begin{tabular}{lcccccc}
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
@@
 \begin{figure}
-	\centering
+	\centering\scalebox{.8}{
 <<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=2.5, fig.width=6>>=
@@
 @
+}
 	\caption{Posterior density curves for each of the three variables (ZINB model). The flood posterior is shown in blue and the tornado posterior is shown in black.}
@@
 \label{s:discussion}
-The goal of this analysis was to determine which of the four models was the best fit for the data.
-In order to make this determination the DIC was calculated for each model on both tornados and flash floods.
-The results are summarized in Table~\ref{t:evalresults}.
-
 \begin{table}
@@
 \label{t:evalresults}
+\scalebox{.8}{
 \begin{tabular}{lrr}
@@
 \bottomrule
-\end{tabular}
+\end{tabular}}
 \end{table}
+The goal of this analysis was to determine which of the four models best fit the data.
+In order to make this determination the DIC was calculated for each model on both tornados and flash floods.
+The results are summarized in Table~\ref{t:evalresults}.
+The Zero Inflated Poisson model has the lowest DIC, for Flash Floods and the ZINB has the lowest for Tornados.
+For estimating or predicting lethality counts for weather it appears that zero inflated models out perform their plain variants however, which zero inflated model is not clear, as we only did two weather events and the ZINB and ZIP were similar.
+When attempting a regression model of death counts from weather events, researchers should consider a zero inflated model.
 
-In general, models with smaller DIC are preferred to models with larger DIC.
-From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados.
-Thus for estimating the rate of leathality of Flash Floods, Tornados, and possibly other severe weather, we should use a Zero Inflated Poisson model to fit the data.
+Negative binomial models tend to perform better than Poisson models when the mean is not equal to the varaince.
+It may be that this was true for the Tornadoes, but not true for the floods.
+The overdispersion of the Tornadoes can be clearly seen when comparing the posterior $\lambda$s for the ZIP model in Figure~\ref{f:zipfig}.
 
-\clearpage
 \begin{thebibliography}{2}

### 2017-05-11 presentation.tex
under length! Yay!

@@
 		\frametitle{Zero Inflated Negative Binomial Model:Results}
+			$2030000$ samples were drawn for each model, with a burn in of $30000$ and one of every $250$ samples was retained, resulting in a final sample size of $8000$.
 		\begin{itemize}
@@
 			\centering
-			\scalebox{0.6}{
+			\scalebox{.6}{
 			\begin{tabular}{lcccccc}
@@
 				\midrule
-				$\sigma$ & 0.307 &
-				0.282&
-				(0.013, 
-				0.716)&
-				0.299 &
-				0.276&
+				$\sigma$ & 0.308 &
+				0.287&
 				(0.012, 
-				0.702)\\
-				$p$ &  0.554 &
+				0.715)&
+				0.29 &
+				0.262&
+				(0.012, 
+				0.693)\\
+				$p$ &  0.555 &
 				0.555&
 				(0.518, 
-				0.589)&
-				0.862 &
+				0.591)&
+				0.863 &
 				0.863&
-				(0.841, 
+				(0.842, 
 				0.882)\\
@@
 				(0.014, 
-				0.052)&
+				0.053)&
 				0.015 &
@@
 				(0.009, 
-				0.033)\\
+				0.031)\\
 				\bottomrule
@@
 			\centering
-			\includegraphics[width=.9\textwidth]{figure/zinbdensity-1} 
-			\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
+			\includegraphics[width=.8\textwidth]{figure/zinbdensity-1} 
 		\end{figure} 
+		\begin{itemize}
+			\item{$p$ can be understood to be ``dangerousness''}
+			\item{$r$ can be thought of as ``duration''}
+		\end{itemize}
+		In this light results make sense: floods are longer lasting (on average) than tornadoes, but less dangerous (on average).
 	\end{frame}
@@
 	\centering
-\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
 	\label{t:evalresults}
-\begin{tabular}{lcc}
+	\begin{tabular}{lrr}
 		\toprule
@@
 		ZIP & 10454 & 8727 \\
-ZINB & 10804.104&7166.309\\
+		ZINB & 10804&7166\\
 		\bottomrule
@@
 
-Models with smaller DIC are preferred.
-From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados.
-Thus for estimating the rate of lethality of Flash Floods, Tornados, we should use a Zero Inflated Poisson model.
+The Zero Inflated Poisson model has the lowest DIC, for Flash Floods and the ZINB has the lowest for Tornadoes.
+For estimating or predicting lethality counts for weather it appears that Zero inflated models out perform their plain variants however, which zero inflated model is not clear, as we only did two weather events and the ZINB and ZIP were similar.
 	\end{frame}
@@
 			\item{See if the results hold on other extreme weather events.}
-			\item{Attempt to fit all data with a mixture model.}
 		\end{enumerate}
